---
title: "descriptive statistics of children (AGE <= 16)"
subtitle: "descr.stat.children.Rmd"
output: html_document
date: "`r format(Sys.time(), '%d %B, %Y')`"
---
  
# setup
```{r, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 7
  )

source("z_start.R") # initalize netezza DB, load core data
if(!exists("TABLE04SAS7BDAT.sparse")) load("./data/TABLE04SAS7BDAT.sparse.RData")
## re-compute missing IXDAYS for all observations
PATIENTS_AGE_REGION <- TABLE04SAS7BDAT.sparse
newIXDAYS                  <- PATIENTS_AGE_REGION$STARTDT - PATIENTS_AGE_REGION$INDEXDT
PATIENTS_AGE_REGION$IXDAYS <- as.numeric(newIXDAYS)

```

# filter PATIENTS 
* AGE <= 16
* 2y MH
* AED treatment
* non-negative DAYSUPP

```{r, include = FALSE}

nPatObsCounter <- function(y) {
  list(addCount=
         function(x,descr) {
           nAdult <- x %>% filter(AGE >= 18) %>% select(ENROLID) %>% distinct() %>% tally()
           if(is.na(nAdult$n)) nAdult <- 0
           nChild <- x %>% filter(AGE <= 16) %>% select(ENROLID) %>% distinct() %>% tally()
           if(is.na(nChild$n)) nChild <- 0
           nObs   <- nrow(x)
           xx <- cbind(descr, nObs, nAdult, nChild)
           colnames(xx) <- c("description","nObs", "Adult, >= 18", "Child, <= 16")
           y <<- rbind(y,xx)
           },
       value = function() return(y)
  )
}

source("functions/n_patient_per_period.R")
source("functions/patient_monthly_enrolment.R")
source("functions/patient_nonNegative_DAYSUPP.R")
source("functions/setAEDFlag.R")
source("functions/aggregate_AED_substancenames.R")
source("functions/add_Phew_PhewHighlevel.R")

# add information
##################
# set AED flag
PATIENTS_AGE_REGION <- setAEDFlag(PATIENTS_AGE_REGION)
# Add Age and Regional Information
PATIENTS_AGE_REGION <- addAgeRegion(PATIENTS_AGE_REGION)
# match PHEWAS terms to ICD9 terms
PATIENTS_AGE_REGION <- add_Phew_PhewHighlevel(PATIENTS_AGE_REGION)
# health insurance class
load("data/monthly_enrolment.RData")
monthly_enrolment <-  
  monthly_enrolment %>% 
  select(ENROLID, db) %>% 
  distinct() %>% 
  mutate(insur = ifelse(db == "CCAE/MDCR", "premium", "basic")) %>% 
  select(-db) %>% 
  as_tibble()
monthly_enrolment$ENROLID <- as.numeric(monthly_enrolment$ENROLID)
PATIENTS_AGE_REGION <- left_join(PATIENTS_AGE_REGION,monthly_enrolment)

# filter
########
patObsCount <- nPatObsCounter(NULL)
patObsCount$addCount(PATIENTS_AGE_REGION,"unconstraint")
# Keep only patients within valid time period 
PATIENTS_AGE_REGION <- filter_patient_medHist_follUp(PATIENTS_AGE_REGION,lbound=-2*365,ubound = 0*365)
patObsCount$addCount(PATIENTS_AGE_REGION,"2y MH")
# only patients with non-negative DAYSUP
PATIENTS_AGE_REGION <- patient_nonNegative_DAYSUPP(PATIENTS_AGE_REGION)
patObsCount$addCount(PATIENTS_AGE_REGION,"non-negative DAYSUPP")
# adult only (AGE <= 16)
PATIENTS_AGE_REGION <- PATIENTS_AGE_REGION %>% filter(AGE <= 16) %>% distinct()
patObsCount$addCount(PATIENTS_AGE_REGION,"AGE <= 16")
# only patients which were treated with AEDs
pat.with.aeds <- PATIENTS_AGE_REGION %>% filter(isAED == TRUE) %>% select(ENROLID) %>% distinct()
PATIENTS_AGE_REGION <- PATIENTS_AGE_REGION %>% filter(ENROLID %in% pat.with.aeds$ENROLID)
patObsCount$addCount(PATIENTS_AGE_REGION,"AED treatment")
#
```

# setup global variables
```{r, include = FALSE}
EPILEPSY <- c("Convulsions", "Epilepsy, recurrent seizures, convulsions",  "Partial epilepsy", "Epilepsy", "Generalized convulsive epilepsy")

```

# get PHEWAS_STRINGS for MAIN.COMORB Strings
1. matching table: main-comorbidity <-> PHEWAS_STRING
2. add main-comorbidity to patients table
```{r, echo = FALSE}

MAIN.COMORB <- c("ADHD","Autism","Migraine","Developmental.delays.and.disorders","Anxiety","Depression")
MAIN.COMORB.PHEWAS <- vector("list",length(MAIN.COMORB))
names(MAIN.COMORB.PHEWAS) <- MAIN.COMORB
MAIN.COMORB.PHEWAS$ADHD <- c("Attention deficit hyperactivity disorder")
MAIN.COMORB.PHEWAS$Autism <- c("Autism")
MAIN.COMORB.PHEWAS$Migraine <- c("Migraine","Migrain with aura")
MAIN.COMORB.PHEWAS$Developmental.delays.and.disorders <- c("Develomental delays and disorders", "Lack of normal physiological development")
MAIN.COMORB.PHEWAS$Anxiety <- c("Anxiety","Anxiety, phobic and dissociative disorders","Generalized anxiety disorder","Agorophobia, social phobia, and panic disorder")
MAIN.COMORB.PHEWAS$Depression <- c("Depression","Major depressive disorder")

# comorb-list to df, join PHEWAS frequency
##########################################
MAIN.COMORB.PHEWAS <-
  data.frame (cbind(
    MAIN.COMORB = as.character(unlist(mapply(rep, MAIN.COMORB, unlist(lapply( MAIN.COMORB.PHEWAS, length))))),
    PHEWAS_STRING = as.character(unlist(MAIN.COMORB.PHEWAS))
  ), row.names = NULL, stringsAsFactors = FALSE)

# 2. add main-comorbidity to patients table
###########################################
PATIENTS_AGE_REGION <- left_join(PATIENTS_AGE_REGION, MAIN.COMORB.PHEWAS)
```


```{r}
# pdf(file="output/descr.stat.Rmd.pdf")
```


# ECDF of first occurrence of a main-comorbidity
```{r}
main.comorb.first.ixday <- PATIENTS_AGE_REGION %>% filter(MAIN.COMORB != "") %>% select(ENROLID, IXDAYS, PHEWAS_STRING, MAIN.COMORB) %>% distinct()
main.comorb.first.ixday <- aggregate(IXDAYS ~., data = main.comorb.first.ixday, FUN = min)
glimpse(main.comorb.first.ixday)

ggplot(main.comorb.first.ixday, aes(IXDAYS)) +  facet_wrap( ~ MAIN.COMORB) + stat_ecdf() + 
  ggtitle(paste("ECDF of first occurrence of diagnosis (PHEWAS) \n aggregated by  main-comorbidity"))

```

# MAIN.COMORB
## prevalence
### by PHEWAS_STRING
```{r, warning=FALSE}

# without any main-comorbidity
x <- PATIENTS_AGE_REGION %>% filter(!is.na(MAIN.COMORB)) %>% select(ENROLID) %>% distinct()
patObsCount$addCount(PATIENTS_AGE_REGION %>% filter(!(ENROLID %in% x$ENROLID)),paste("without any main-comorbidity"))

# with main co-morbidities
###############################
buffer <- 180 # 3 months + 3 months
x <- main.comorb.first.ixday %>% filter(IXDAYS >= buffer) %>% select(ENROLID) %>% distinct()
patObsCount$addCount(PATIENTS_AGE_REGION %>% filter(ENROLID %in% x$ENROLID),paste("incident  main-comorbidities >= ",buffer, " days"))


# n patients per MAIN.COMORB, by PHEWAS_STRING
##############################################
MAIN.COMORB.PHEWAS <-
  left_join(
  MAIN.COMORB.PHEWAS,
  PATIENTS_AGE_REGION %>%
  filter(PHEWAS_STRING %in% MAIN.COMORB.PHEWAS$PHEWAS_STRING) %>%
  select(ENROLID, PHEWAS_STRING) %>%
  distinct() %>%
  group_by(PHEWAS_STRING) %>%
  tally(),
  by = "PHEWAS_STRING"
  )

if(any(is.na(MAIN.COMORB.PHEWAS$n))) MAIN.COMORB.PHEWAS$n[is.na(MAIN.COMORB.PHEWAS$n)] <- 0
knitr::kable(MAIN.COMORB.PHEWAS)
```

### by MAIN.COMORB
```{r, echo = FALSE}
# count prevalent patients per MC
prev.allAEDs <- 
  PATIENTS_AGE_REGION %>% 
  filter(MAIN.COMORB != "") %>% 
  select(ENROLID, MAIN.COMORB) %>% 
  distinct() %>% 
  group_by(MAIN.COMORB) %>% 
  tally() %>% 
  dplyr::rename(n.prevalent = n)

knitr::kable(prev.allAEDs[order(-prev.allAEDs$n.prevalent),])

cat("The sum of patients within a comorbidity is NOT the sum of patients in the PHEWAS_STRING by MAIN.COMORB due to overlaps. This approach leads to an overestimation")
```

### incidence
```{r, echo = FALSE}
# 2. plot number of patients with incident main-comorbidity occurrence as a function of <buffer.int>-time
#########################################################################################################
buffer.int <- 1:365

# summarised over all patients
##############################
x <- unlist(mapply(function(x) main.comorb.first.ixday %>% filter(IXDAYS >= x) %>% select(ENROLID) %>% distinct() %>% tally(), buffer.int))
x <- as.data.frame(cbind(buffer.int, x))

ggplot(x, aes(buffer.int, x)) +
  geom_line(aes(y = x, colour = "sum") ) +
  ggtitle("Number of patients with new  main-comorbidities at time >= IXDAYS \n summarized over all  main-comorbidities") +
  xlab("IXDAYS")

# by main-comorbidity
#####################
first.ixdays.vs.buffer <- as.data.frame(matrix(data = NA, nrow = length(buffer.int), ncol = length(MAIN.COMORB)))
colnames(first.ixdays.vs.buffer) <- MAIN.COMORB
first.ixdays.vs.buffer <- cbind(buffer = buffer.int, first.ixdays.vs.buffer)

for( m in unique(main.comorb.first.ixday$MAIN.COMORB)) {
  first.ixdays.vs.buffer[, colnames(first.ixdays.vs.buffer) == m] <-
    unlist(
      mapply(
        function(x)
          main.comorb.first.ixday %>% filter(MAIN.COMORB == m, IXDAYS >= x) %>% select(ENROLID) %>% distinct() %>% tally(),
        x = buffer.int))
}

ggplot(first.ixdays.vs.buffer, aes(buffer)) +
  geom_line(aes(y = ADHD, colour = "ADHD") ) +
  geom_line(aes(y= Autism, colour = "Autism")) +
  geom_line(aes(y= Migraine, colour =  "Migraine")) +
  geom_line(aes(y= Developmental.delays.and.disorders, colour = "Developmental delays and disorders")) +
  geom_line(aes(y= Anxiety, colour = "Anxiety")) +
  geom_line(aes(y= Depression, colour = "Depression")) +
  ggtitle("Number of patients with new  main-comorbidities at time >= IXDAYS \n by  main-comorbidity") +
  xlab("IXDAYS")

mc.inc180 <- first.ixdays.vs.buffer %>% filter(buffer == 180) %>% dplyr::select(-buffer) %>% t() 
mc.inc180 <- data.frame(MC = row.names(mc.inc180), n.incident = mc.inc180)
mc.inc180 <- mc.inc180[order(-mc.inc180$n.incident),]
row.names(mc.inc180) <- c()
knitr::kable(mc.inc180)
```


```{r}
# dev.off()
```

# patient frequency by filter criterium
```{r}

ggplot(gather(patObsCount$value()[-2],"age","n",2:3), aes(x=description,y=n, fill=age, label = n)) +
  geom_bar(aes(fill=age), position="dodge",stat="identity") +
  geom_text(aes(label=n), position=position_dodge(0.9), vjust=-1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  ggtitle("patient frequency by filter criterium and by age")

knitr::kable(patObsCount$value())
```

# descriptive stats
* age histogram
* AED histogram
* non-AED histogram, top 20
* diagnosis histogram (by PHEWAS, PHEWAS_highlevel)


*The Truven Health MarketScan® Databases Insurance classes:*

>The Truven Health MarketScan Databases consist of several different datasets. The Commercial Claims and Encounters (CCAE) database is a nationally representative collection of inpatient, outpatient, and pharmaceutical claims from more than 200 insurance carriers and large, self-insuring companies. The Medicare Supplemental and Coordination of Benefits (MDCR) database consists of claims for Medicare-eligible retirees with employer-sponsored supplemental plans.


```{r}
# age 
######
# age histogram
ggplot(PATIENTS_AGE_REGION %>% select(ENROLID, AGE) %>% distinct(), aes(AGE)) +
  ggtitle(paste("AGE histogram","\n total pop.-size = ",PATIENTS_AGE_REGION %>% select(ENROLID) %>% distinct() %>% tally() )) +
  geom_histogram(aes(y = ..density..))

# age histogram by health insurance
ggplot(PATIENTS_AGE_REGION %>% select(insur, ENROLID, AGE) %>% distinct(), aes(AGE)) +
  facet_wrap(~insur) +
  ggtitle(paste("AGE histogram, by insurance flag","\n total pop.-size = ",PATIENTS_AGE_REGION %>% select(ENROLID) %>% distinct() %>% tally() )) +
  geom_histogram(aes(y = ..density..))

ggplot(PATIENTS_AGE_REGION %>% select(insur, ENROLID, AGE) %>% distinct(), aes(AGE)) +
  facet_wrap(~insur) +
  ggtitle(paste("AGE histogram, by insurance flag","\n total pop.-size = ",PATIENTS_AGE_REGION %>% select(ENROLID) %>% distinct() %>% tally() )) +
  geom_histogram()

# PHARMCLS1 histogram
#####################

# PHARMCLS1 histogram 
ggplot(
    PATIENTS_AGE_REGION %>% 
    filter(SUBSTANCENAME != "", PHARMCLS1 != "") %>% 
    select(PHARMCLS1, isAED, ENROLID) %>% 
    distinct() %>% 
    group_by(isAED, PHARMCLS1) %>% 
    tally(sort=TRUE) %>% 
    top_n(20,n),
  aes(x = reorder(PHARMCLS1, n), y = n, fill= isAED)
  ) +
  coord_flip() +
  geom_bar(stat="identity") +
  ggtitle(paste("PHARMCLS1 frequency, top 2x20","\n total pop.-size = ",PATIENTS_AGE_REGION %>% select(ENROLID) %>% distinct() %>% tally() ))

# SUBSTANCE frequencies
ggplot(
  PATIENTS_AGE_REGION %>%
    filter(SUBSTANCENAME != "") %>% 
    select(ENROLID, isAED, SUBSTANCENAME) %>%
    distinct() %>%
    group_by(isAED, SUBSTANCENAME) %>%
    tally(sort = TRUE) %>%
    top_n(20, n) %>% 
    filter(n >= 100),
  aes(x = reorder(SUBSTANCENAME, n), y = n)) +
  facet_wrap( ~ isAED) +
  coord_flip() +
  geom_bar(stat = "identity") +
  ggtitle(
    paste(
      "SUBSTANCE frequency, top 20",
      "\nnon-AED vs AED",
      "\n total pop.-size = ",
      PATIENTS_AGE_REGION %>% select(ENROLID) %>% distinct() %>% tally())
  ) 

# DIAGNOSIS by PHEWAS_STRING
ggplot(
  PATIENTS_AGE_REGION %>%
    filter(PHEWAS_STRING != "") %>% 
    select(ENROLID, MAIN.COMORB, PHEWAS_STRING) %>%
    distinct() %>%
    group_by( MAIN.COMORB, PHEWAS_STRING) %>%
    tally(sort = TRUE) %>%
    top_n(20, n),
  aes(x = reorder(PHEWAS_STRING, n), y = n, fill = MAIN.COMORB)) +
  coord_flip() +
  geom_bar(stat = "identity") +
  ggtitle(
    paste(
      "PHEWAS frequency, top 20 (each category)",
      "\n total pop.-size = ",
      PATIENTS_AGE_REGION %>% select(ENROLID) %>% distinct() %>% tally())
  ) 

```

<!-- # prevalence vs incidence (>= +180 days) -->

<!-- * for the top 20 co-morbidities -->
<!-- ```{r, warning=FALSE} -->
<!-- top20.cm <-  -->
<!--   PATIENTS_AGE_REGION %>% -->
<!--   filter(PHEWAS_STRING != "") %>%  -->
<!--   select(ENROLID, PHEWAS_STRING) %>% -->
<!--   distinct() %>% -->
<!--   group_by( PHEWAS_STRING) %>% -->
<!--   tally(sort = TRUE) %>% -->
<!--   top_n(20, n) -->

<!-- top20.inc <-  -->
<!--   PATIENTS_AGE_REGION %>% -->
<!--   filter(PHEWAS_STRING %in% top20.cm$PHEWAS_STRING) %>%  -->
<!--   select(ENROLID, PHEWAS_STRING, IXDAYS) %>% -->
<!--   distinct()  -->

<!-- top20.inc <- -->
<!--   aggregate(formula = IXDAYS ~ ENROLID + PHEWAS_STRING, data = top20.inc, FUN = min) %>%  -->
<!--   filter(IXDAYS >= 180) %>%  -->
<!--   group_by(PHEWAS_STRING) %>%  -->
<!--   tally(sort = TRUE) -->


<!-- knitr::kable( -->
<!--   left_join(top20.cm %>% rename(n.prevalence = n), -->
<!--           top20.inc %>% rename(n.incidence = n), -->
<!--           by = "PHEWAS_STRING") %>%  -->
<!--   mutate(perc.incidence = round(n.incidence / n.prevalence,2)) -->
<!--           ) -->

<!-- ``` -->

