---
title: "descriptive statistics of adult (AGE >= 18)"
output: html_notebook
date: "2017 JAN 18"
---

# setup global variables
```{r, echo = FALSE, include = FALSE}
source("z_start.R") # initalize netezza DB, load core data

params <- vector("list")
params$incidence.buffer <- 180
params$EPILEPSY <-
  c(
  "Convulsions",
  "Epilepsy, recurrent seizures, convulsions",
  "Partial epilepsy",
  "Epilepsy",
  "Generalized convulsive epilepsy"
  )

```  
 
# setup
```{r, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 7
  )

if(!exists("TABLE04SAS7BDAT")) load("./data/TABLE04SAS7BDAT.RData")

PATIENTS_AGE_REGION <- 
  TABLE04SAS7BDAT %>% 
  select(-EVTDATE, -MHSAFL, -HOSPFL, -NDCNUM, -(PHARMCLS2:PHARMCLS5), - THERCLS, -THERGRP)

newIXDAYS                  <- PATIENTS_AGE_REGION$STARTDT - PATIENTS_AGE_REGION$INDEXDT
PATIENTS_AGE_REGION$IXDAYS <- as.numeric(newIXDAYS)
rm(newIXDAYS)

```

# filter PATIENTS 
* AGE >= 18
* 2y MH
* AED treatment
* non-negative DAYSUPP

```{r, include = FALSE}

nPatObsCounter <- function(y) {
  list(addCount=
         function(x,descr) {
           nAdult <- x %>% filter(AGE >= 18) %>% select(ENROLID) %>% distinct() %>% tally()
           nChild <- x %>% filter(AGE <= 16) %>% select(ENROLID) %>% distinct() %>% tally()
           nObs   <- nrow(x)
           xx <- cbind(descr, nObs, nAdult, nChild)
           colnames(xx) <- c("description","nObs", "Adult, >= 18", "Child, <= 16")
           y <<- rbind(y,xx)
           },
       value = function() return(y)
  )
}

source("functions/n_patient_per_period.R")
source("functions/patient_monthly_enrolment.R")
source("functions/patient_nonNegative_DAYSUPP.R")
source("functions/setAEDFlag.R")
source("functions/aggregate_AED_substancenames.R")
source("functions/add_Phew_PhewHighlevel.R")

# add information
##################
# set AED flag
PATIENTS_AGE_REGION <- setAEDFlag(PATIENTS_AGE_REGION)
# Add Age and Regional Information
PATIENTS_AGE_REGION <- addAgeRegion(PATIENTS_AGE_REGION)
# match PHEWAS terms to ICD9 terms
PATIENTS_AGE_REGION <- add_Phew_PhewHighlevel(PATIENTS_AGE_REGION)
# health insurance class
load("data/monthly_enrolment.RData")
monthly_enrolment <-  
  monthly_enrolment %>% 
  select(ENROLID, db) %>% 
  distinct() %>% 
  mutate(insur = ifelse(db == "CCAE/MDCR", "premium", "basic")) %>% 
  select(-db) %>% 
  as_tibble()
monthly_enrolment$ENROLID <- as.numeric(monthly_enrolment$ENROLID)
PATIENTS_AGE_REGION <- left_join(PATIENTS_AGE_REGION,monthly_enrolment)

# filter
########
patObsCount <- nPatObsCounter(NULL)
patObsCount$addCount(PATIENTS_AGE_REGION,"unconstraint")

# Keep only patients within valid time period 
PATIENTS_AGE_REGION <- filter_patient_medHist_follUp(PATIENTS_AGE_REGION,lbound=-2*365,ubound = 0*365)
patObsCount$addCount(PATIENTS_AGE_REGION,"2y MH")

# only patients with non-negative DAYSUP
PATIENTS_AGE_REGION <- patient_nonNegative_DAYSUPP(PATIENTS_AGE_REGION)
patObsCount$addCount(PATIENTS_AGE_REGION,"non-negative DAYSUPP")

# adult only (AGE >= 18)
PATIENTS_AGE_REGION <- PATIENTS_AGE_REGION %>% filter(AGE >= 18) %>% distinct()
patObsCount$addCount(PATIENTS_AGE_REGION,"AGE >= 18")

# only patients which were treated with AEDs
pat.with.aeds <- 
  PATIENTS_AGE_REGION %>%
  filter(isAED == TRUE) %>% 
  select(ENROLID) %>% 
  distinct()

PATIENTS_AGE_REGION <- 
  PATIENTS_AGE_REGION %>% 
  filter(ENROLID %in% pat.with.aeds$ENROLID)

patObsCount$addCount(PATIENTS_AGE_REGION,"AED treatment")

# patch ICD9 Code V12.54
## add main comorbidities
PATIENTS_AGE_REGION <- AddMainComorb(PATIENTS_AGE_REGION)
## Patch ICD9 Code V12.54
source("functions/Patch_V12.54.R")
PATIENTS_AGE_REGION <- Patch_V12.54(PATIENTS_AGE_REGION)

#
PATIENTS_AGE_REGION <- unique(PATIENTS_AGE_REGION)
```

# get PHEWAS_STRINGS for MAIN.COMORB Strings
1. matching table: main-comorbidity <-> PHEWAS_STRING
2. add main-comorbidity to patients table
```{r}
# list of PHEWAS_STRINGS with their assignments to co-morbidiets
  MAIN.COMORB <- c("Hypertension","Diabetes","Hyperlipidemia","Anxiety","Migraine","Depression","Stroke.IschemAttack")
  MAIN.COMORB.PHEWAS <- vector("list",length(MAIN.COMORB))
  names(MAIN.COMORB.PHEWAS) <- MAIN.COMORB
  MAIN.COMORB.PHEWAS$Hypertension <- c("Essential hypertension")
  MAIN.COMORB.PHEWAS$Diabetes <- c("Type 2 diabetes")
  MAIN.COMORB.PHEWAS$Hyperlipidemia <- c("Hyperlipidemia","Mixed hyperlipidemia")
  MAIN.COMORB.PHEWAS$Anxiety <- c("Anxiety disorder","Generalized anxiety disorder","Anxiety, phobic and dissociative disorders","Agorophobia, social phobia, and panic disorder")
  MAIN.COMORB.PHEWAS$Stroke.IschemAttack <- c("Ischemic stroke", "Transient cerebral ischemia")
  MAIN.COMORB.PHEWAS$Migraine <- c("Migraine","Migrain with aura")
  MAIN.COMORB.PHEWAS$Depression <- c("Depression","Major depressive disorder")
  
  # comorb-list to df, join PHEWAS frequency
  MAIN.COMORB.PHEWAS <-
    data.frame (cbind(
      MAIN.COMORB = as.character(unlist(mapply(rep, MAIN.COMORB, unlist(lapply( MAIN.COMORB.PHEWAS, length))))),
      PHEWAS_STRING = as.character(unlist(MAIN.COMORB.PHEWAS))
    ), row.names = NULL, stringsAsFactors = FALSE)
  
  knitr::kable(MAIN.COMORB.PHEWAS)
  
  # add main-comorbidity to patients table
  PATIENTS_AGE_REGION <- left_join(PATIENTS_AGE_REGION, MAIN.COMORB.PHEWAS)
```


```{r}
# pdf(file="output/descr.stat.Rmd.pdf")
```


# ECDF of first occurrence of a main-comorbidity
```{r}
main.comorb.first.ixday <-
  PATIENTS_AGE_REGION %>% 
  filter(MAIN.COMORB != "") %>% 
  select(ENROLID, IXDAYS, MAIN.COMORB) %>% 
  distinct()

main.comorb.first.ixday <- aggregate(IXDAYS ~., data = main.comorb.first.ixday , FUN = min)
glimpse(main.comorb.first.ixday)

ggplot(main.comorb.first.ixday, aes(IXDAYS)) +  
  facet_wrap( ~ MAIN.COMORB) +
  stat_ecdf() + 
  ggtitle(paste("ECDF of first occurrence of diagnosis (PHEWAS) \n aggregated by main-comorbidity \n blue line = 180 days")) +
  geom_vline(xintercept = 180, colour = "blue", linetype = "longdash")

```

# MAIN.COMORB
## prevalence
### by PHEWAS_STRING
```{r, warning=FALSE}

# without any main-comorbidity
x <- 
  PATIENTS_AGE_REGION %>% 
  filter(!is.na(MAIN.COMORB)) %>% 
  select(ENROLID) %>%
  distinct()

patObsCount$addCount(PATIENTS_AGE_REGION %>% filter(!(ENROLID %in% x$ENROLID)),paste("without any main-comorbidity"))

# with main co-morbidities
x <- 
  main.comorb.first.ixday %>% 
  filter(IXDAYS >= params$incidence.buffer) %>% 
  select(ENROLID) %>%
  distinct()

patObsCount$addCount(PATIENTS_AGE_REGION %>% filter(ENROLID %in% x$ENROLID),paste("incident main-comorbidities >= ",params$incidence.buffer, " days"))


# n patients per MAIN.COMORB, by PHEWAS_STRING
##############################################
MAIN.COMORB.PHEWAS <-
  left_join(
  MAIN.COMORB.PHEWAS,
  PATIENTS_AGE_REGION %>%
  filter(PHEWAS_STRING %in% MAIN.COMORB.PHEWAS$PHEWAS_STRING) %>%
  select(ENROLID, PHEWAS_STRING) %>%
  distinct() %>%
  group_by(PHEWAS_STRING) %>%
  tally(),
  by = "PHEWAS_STRING"
  )

knitr::kable(MAIN.COMORB.PHEWAS)
```

### by MAIN.COMORB
```{r, echo = FALSE}
# count prevalent patients per MC
prev.allAEDs <- 
  PATIENTS_AGE_REGION %>% 
  filter(MAIN.COMORB != "") %>% 
  select(ENROLID, MAIN.COMORB) %>% 
  distinct() %>% 
  group_by(MAIN.COMORB) %>% 
  tally() %>% 
  dplyr::rename(n.prevalent = n)

knitr::kable(prev.allAEDs[order(-prev.allAEDs$n.prevalent),])

cat("The sum of patients within a comorbidity is NOT the sum of patients in the PHEWAS_STRING by MAIN.COMORB due to overlaps. This approach leads to an overestimation")
```

## incidence
```{r, warning=FALSE}
# 2. plot number of patients with incident main-comorbidity occurrence as a function of <buffer.int>-time
#########################################################################################################
buffer.int <- 1:365

# summarised over all patients
##############################
x <- 
  unlist(
    mapply(
      function(x) main.comorb.first.ixday %>% filter(IXDAYS >= x) %>% select(ENROLID) %>% distinct() %>% tally(),
      buffer.int)
    )

x <- as.data.frame(cbind(buffer.int, x))

ggplot(x, aes(buffer.int, x)) +
  geom_line(aes(y = x, colour = "sum") ) +
  ggtitle("Number of patients with new main-comorbidities at time >= IXDAYS \n summarized over all main-comorbidities") +
  xlab("IXDAYS")

# by main-comorbidity
#####################
first.ixdays.vs.buffer <- as.data.frame(matrix(data = NA, nrow = length(buffer.int), ncol = length(MAIN.COMORB)))
colnames(first.ixdays.vs.buffer) <- MAIN.COMORB
first.ixdays.vs.buffer <- cbind(buffer = buffer.int, first.ixdays.vs.buffer)

for( m in unique(main.comorb.first.ixday$MAIN.COMORB)) {
  first.ixdays.vs.buffer[, colnames(first.ixdays.vs.buffer) == m] <-
    unlist(
      mapply(
        function(x)
          main.comorb.first.ixday %>% filter(MAIN.COMORB == m, IXDAYS >= x) %>% select(ENROLID) %>% distinct() %>% tally(),
        x = buffer.int))
}

ggplot(first.ixdays.vs.buffer, aes(buffer.int)) +
  geom_line(aes(y = Hypertension, colour = "Hypertension") ) +
  geom_line(aes(y= Diabetes, colour = "Diabetes")) +
  geom_line(aes(y= Hyperlipidemia, colour =  "Hyperlipidemia")) +
  geom_line(aes(y= Anxiety, colour = "Anxiety")) + 
  geom_line(aes(y= Stroke.IschemAttack, colour = "Stroke, ischemic attack")) +
  geom_line(aes(y= Migraine, colour = "Migraine")) +
  geom_line(aes(y= Depression, colour = "Depression")) +
  ggtitle("Number of patients with new main-comorbidities at time >= IXDAYS \n by main-comorbidity") +
  xlab("IXDAYS")

mc.inc180 <- 
  first.ixdays.vs.buffer %>%
  filter(buffer == 180) %>% 
  dplyr::select(-buffer) %>%
  t() 

mc.inc180 <- data.frame(MC = row.names(mc.inc180), n.incident = mc.inc180)
mc.inc180 <- mc.inc180[order(-mc.inc180$n.incident),]
row.names(mc.inc180) <- c()
knitr::kable(mc.inc180 )
```

# patient frequency by filter criterium
```{r}

ggplot(gather(patObsCount$value()[-2],"age","n",2:3), aes(x=description,y=n, fill=age, label = n)) +
  geom_bar(aes(fill=age), position="dodge",stat="identity") +
  geom_text(aes(label=n), position=position_dodge(0.9), vjust=-1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  ggtitle("patient frequency by filter criterium and by age")

knitr::kable(patObsCount$value())
```

```{r}
# dev.off()
```

# descriptive stats
* age histogram
* AED histogram
* non-AED histogram, top 20
* diagnosis histogram (by PHEWAS, PHEWAS_highlevel)

*The Truven Health MarketScan® Databases Insurance classes:*

>The Truven Health MarketScan Databases consist of several different datasets. The Commercial Claims and Encounters (CCAE) database is a nationally representative collection of inpatient, outpatient, and pharmaceutical claims from more than 200 insurance carriers and large, self-insuring companies. The Medicare Supplemental and Coordination of Benefits (MDCR) database consists of claims for Medicare-eligible retirees with employer-sponsored supplemental plans.

```{r}
# age 
######

# age histogram
ggplot(PATIENTS_AGE_REGION %>% select(ENROLID, AGE) %>% distinct(), aes(AGE)) +
  ggtitle(paste("AGE histogram","\n total pop.-size = ",PATIENTS_AGE_REGION %>% select(ENROLID) %>% distinct() %>% tally() )) +
  geom_histogram(aes(y = ..density..))

# age histogram by health insurance
ggplot(PATIENTS_AGE_REGION %>% select(insur, ENROLID, AGE) %>% distinct(), aes(AGE)) +
  facet_wrap(~insur) +
  ggtitle(paste("AGE histogram, by insurance flag","\n total pop.-size = ",PATIENTS_AGE_REGION %>% select(ENROLID) %>% distinct() %>% tally() )) +
  geom_histogram(aes(y = ..density..))

ggplot(PATIENTS_AGE_REGION %>% select(insur, ENROLID, AGE) %>% distinct(), aes(AGE)) +
  facet_wrap(~insur) +
  ggtitle(paste("AGE histogram, by insurance flag","\n total pop.-size = ",PATIENTS_AGE_REGION %>% select(ENROLID) %>% distinct() %>% tally() )) +
  geom_histogram()

# PHARMCLS1 histogram
#####################

# PHARMCLS1 histogram 
ggplot(
    PATIENTS_AGE_REGION %>% 
    filter(SUBSTANCENAME != "", PHARMCLS1 != "") %>% 
    select(PHARMCLS1, isAED, ENROLID) %>% 
    distinct() %>% 
    group_by(isAED, PHARMCLS1) %>% 
    tally(sort=TRUE) %>% 
    top_n(20,n),
  aes(x = reorder(PHARMCLS1, n), y = n, fill= isAED)
  ) +
  coord_flip() +
  geom_bar(stat="identity") +
  ggtitle(paste("PHARMCLS1 frequency, top 2x20","\n total pop.-size = ",PATIENTS_AGE_REGION %>% select(ENROLID) %>% distinct() %>% tally() ))

# SUBSTANCE frequencies
ggplot(
  PATIENTS_AGE_REGION %>%
    select(ENROLID, isAED, SUBSTANCENAME) %>%
    distinct() %>%
    group_by(isAED, SUBSTANCENAME) %>%
    tally(sort = TRUE) %>%
    top_n(20, n) %>% 
    filter(n >= 100),
  aes(x = reorder(SUBSTANCENAME, n), y = n)) +
  facet_wrap( ~ isAED) +
  coord_flip() +
  geom_bar(stat = "identity") +
  ggtitle(
    paste(
      "SUBSTANCE frequency, top 20",
      "\nnon-AED vs AED",
      "\n total pop.-size = ",
      PATIENTS_AGE_REGION %>% select(ENROLID) %>% distinct() %>% tally())
  ) 

# DIAGNOSIS by PHEWAS_STRING
ggplot(
  PATIENTS_AGE_REGION %>%
    filter(PHEWAS_STRING != "") %>% 
    select(MAIN.COMORB, ENROLID, PHEWAS_STRING) %>%
    distinct() %>%
    group_by(MAIN.COMORB, PHEWAS_STRING) %>%
    tally(sort = TRUE) %>%
    top_n(20, n),
  aes(x = reorder(PHEWAS_STRING, n), y = n, fill = MAIN.COMORB)) +
  coord_flip() +
  geom_bar(stat = "identity") +
  ggtitle(
    paste(
      "PHEWAS frequency, top 20 (each category)",
      "\n total pop.-size = ",
      PATIENTS_AGE_REGION %>% select(ENROLID) %>% distinct() %>% tally())
  ) 

```

<!-- # prevalence vs incidence (>= +180 days) -->

<!-- * for the top 20 co-morbidities -->
<!-- ```{r} -->

<!-- top20.cm <-  -->
<!--   PATIENTS_AGE_REGION %>% -->
<!--   filter(PHEWAS_STRING != "") %>%  -->
<!--   select(ENROLID, PHEWAS_STRING) %>% -->
<!--   distinct() %>% -->
<!--   group_by( PHEWAS_STRING) %>% -->
<!--   tally(sort = TRUE) %>% -->
<!--   top_n(20, n) -->

<!-- top20.inc <-  -->
<!--   PATIENTS_AGE_REGION %>% -->
<!--   filter(PHEWAS_STRING %in% top20.cm$PHEWAS_STRING) %>%  -->
<!--   select(ENROLID, PHEWAS_STRING, IXDAYS) %>% -->
<!--   distinct()  -->

<!-- top20.inc <- -->
<!--   aggregate(formula = IXDAYS ~ ENROLID + PHEWAS_STRING, data = top20.inc, FUN = min) %>%  -->
<!--   filter(IXDAYS >= 180) %>%  -->
<!--   group_by(PHEWAS_STRING) %>%  -->
<!--   tally(sort = TRUE) -->


<!-- knitr::kable( -->
<!--   left_join(top20.cm %>% rename(n.prevalence = n), -->
<!--           top20.inc %>% rename(n.incidence = n), -->
<!--           by = "PHEWAS_STRING") %>%  -->
<!--   mutate(perc.incidence = round(n.incidence / n.prevalence,2)) -->
<!--           ) -->
<!-- ``` -->

