---
title: "disease_Descriptors.Rmd"
subtitle: ""
author: "Thomas Gerlach"
date: "2017 JAN 30"
output: html_notebook
---

# setup
```{r, echo=FALSE, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=10, fig.height=7) 
source("z_start.R") # initalize netezza DB, load core data
```

## script specific functions
```{r}


##################
ParseConcatenatedICD9 <- function(icd9.unparsed) {
  
  ICD9GROUPING <- tbl(src,"REFERENCE.ICD9GROUPING") %>% filter(ICD9 != "") %>%   distinct() %>%  collect()
  ICD9GROUPING <- as_tibble(apply(ICD9GROUPING,2,as.character))
  ICD9 <- suppressWarnings(as_tibble(cbind(ICD9c = ICD9GROUPING$ICD9, ICD9n = ICD9GROUPING$ICD9 %>% as.character() %>% as.numeric())))
  ICD9 <- suppressWarnings(data.frame(ICD9c = ICD9GROUPING$ICD9, ICD9n = as.numeric(as.character(ICD9GROUPING$ICD9))))
  
  parsed <- c()
  
  for(k in 1:length(icd9.unparsed)) { 
    
     if(k %% 100 == 0) print(k)
    #print(k)
    x1 <- strsplit(icd9.unparsed[k],",") %>% unlist() %>% strsplit("-")
    x2 <- unlist(lapply(x1,length))
    
    # parse and attach non-interval elements
    x2.nonint <- which(x2==1)
    if(length(x2.nonint)>0){
      x2.nonint <- trimws(unlist(x1[x2.nonint]))
      parsed    <- rbind(parsed,cbind(icd9.unparsed[k],ICD9=x2.nonint,row.names=NULL))
    }
    
    # parse and attach interval-elements (e.g "120-121")
    x2.interv <- which(x2==2)
    
    if(length(x2.interv) >0) {
      
      for (l in 1:length(x2.interv)) {
        x3 <- unlist(x1[[x2.interv[l]]])
        lb <- as.numeric(x3[1])
        rb <- as.numeric(x3[2])
        if(rb == floor(rb)) rb <- rb + 1 - 1e-3
        x3.ICD9.idx <- which(between(ICD9$ICD9n,lb,rb))
        parsed      <- rbind(parsed,cbind(icd9.unparsed[k],ICD9=as.character(ICD9$ICD9c[x3.ICD9.idx]),row.names=NULL))
      }
    }
  }
  colnames(parsed) <- c("unparsed","parsed")
  return(parsed)
}
```


# load + preprocess matching resources
Preprocessing:
* Normalize Strings
* make.names()

## ICD9GROUPING
```{r}
ICD9GROUPING <- tbl(src, "REFERENCE.ICD9GROUPING") %>% filter(ICD9 != "") %>% collect()
ICD9GROUPING <- NormalizeStrings(ICD9GROUPING) %>% as_tibble()
```

todo(check next chunk)

```{r}
DISEASES <- ICD9GROUPING %>% filter(ICD9 != "") %>% dplyr::select(ICD9) %>% distinct()
```


## MESH2ICD9
```{r, echo = FALSE}
MESH2ICD9 <- read.csv("~/epilepsie_preprocessing_descrStat/data/diseaseDescriptors/doid_MESH2ICD9_2016-09-15.csv", header=TRUE, sep=";")
MESH2ICD9 <- as_tibble(unique(MESH2ICD9))

# drop IDs which no MESH terms
MESH2ICD9$mesh[MESH2ICD9$mesh == ""] <- NA
x <- MESH2ICD9 %>% group_by(id) %>% dplyr::select(mesh) %>% filter(!is.na(mesh)) %>% distinct()
MESH2ICD9 <- MESH2ICD9 %>% filter(id %in% x$id)

# drop IDs which no ICD9 terms
x <- MESH2ICD9 %>% group_by(id) %>% dplyr::select(icd9) %>% filter(!is.na(icd9)) %>% distinct()
MESH2ICD9 <- MESH2ICD9 %>% filter(id %in% x$id)

# reshape
MESH2ICD9x <- MESH2ICD9 %>% dplyr::select(id, mesh) %>% filter(!is.na(mesh))
MESH2ICD9x <- left_join(MESH2ICD9x, MESH2ICD9 %>% dplyr::select(id, icd9) %>% filter(!is.na(icd9)))
MESH2ICD9  <- MESH2ICD9x %>% dplyr::select(-id) %>% distinct()
rm(MESH2ICD9x, x)
```

## DISEASE2BIOMARKER
```{r}
DISEASE2BIOMARKER <- tbl(src, "REFERENCE.DISEASE2BIOMARKER" ) %>% collect()
DISEASE2BIOMARKER <- DISEASE2BIOMARKER %>% dplyr::select(-DISEASE, -ICD10, -BIOMARKER_REC_ID) %>% distinct()
DISEASE2BIOMARKER <- apply(DISEASE2BIOMARKER,2,as.character) %>% as_tibble()

# parse ICD9
icd9parsed        <- as_tibble(ParseConcatenatedICD9( as.character(unique(DISEASE2BIOMARKER$ICD9)) ))
DISEASE2BIOMARKER <- 
  left_join(DISEASE2BIOMARKER, icd9parsed, by = c("ICD9" = "unparsed")) %>% 
  dplyr::select(-ICD9) %>%
  dplyr::rename(ICD9 = parsed) %>%
  distinct()
rm(icd9parsed)

DISEASE2BIOMARKER$BIOMARKERNAME <- make.names(trimws(toupper(DISEASE2BIOMARKER$BIOMARKERNAME)))
```

## DISEASE2PHEWAS
```{r}
ICD9GROUPING <- tbl(src,"REFERENCE.ICD9GROUPING") %>% dplyr::select(ICD9, ICD9_STRING, PHEWAS_CODE, PHEWAS_STRING) %>% distinct() %>% collect()
ICD9GROUPING <- NormalizeStrings(ICD9GROUPING)
ICD9GROUPING$ICD9 <- as.character(ICD9GROUPING$ICD9)
ICD9GROUPING$PHEWAS_STRING <- as.character(ICD9GROUPING$PHEWAS_STRING)
ICD9GROUPING$PHEWAS_STRING <- trimws(toupper(ICD9GROUPING$PHEWAS_STRING))

# omit E- and V- ICD9-codings
E <- grep("E",ICD9GROUPING$ICD9)
V <- grep("V",ICD9GROUPING$ICD9)
ICD9GROUPING <- ICD9GROUPING[-c(E,V),]
rm(E,V)

# erase / unclear missing values
ICD9GROUPING <- ICD9GROUPING %>% filter(ICD9 != "", !is.na(PHEWAS_CODE), PHEWAS_STRING != "", PHEWAS_STRING != "1", PHEWAS_STRING != "-")

# PHEWAS_CODE -> PHEWAS_HIGHLEVEL_CODE
# PHEWAS_HIGHLEVEL_CODE -> PHEWAS_HIGHLEVEL_STRING
ICD9GROUPING$PHEWAS_CODE <- as.numeric(ICD9GROUPING$PHEWAS_CODE)

PHEWAS_HL <- 
  ICD9GROUPING %>% 
  mutate(PHEWAS_floor = floor(PHEWAS_CODE)) %>% 
  mutate(isHighlevel = PHEWAS_floor == PHEWAS_CODE) %>% 
  filter(isHighlevel == TRUE) %>% 
  dplyr::select(PHEWAS_CODE, PHEWAS_STRING) %>%
  distinct() %>% 
  dplyr::rename(PHEWAS_CODE_HL = PHEWAS_CODE, PHEWAS_STRING_HL = PHEWAS_STRING)

ICD9GROUPING <- 
  ICD9GROUPING %>% 
  mutate(PHEWAS_CODE_HL = floor(PHEWAS_CODE)) 

# ICD9GROUPING2
ICD9GROUPING <- inner_join(ICD9GROUPING, PHEWAS_HL)

DISEASE2PHEWAS <-
  ICD9GROUPING %>%
  dplyr::select(-ICD9_STRING,-PHEWAS_CODE, -PHEWAS_CODE_HL,-PHEWAS_STRING_HL) %>%
  distinct()

DISEASE2PHEWAS_HL <- 
  ICD9GROUPING %>% 
  dplyr::select(-ICD9_STRING, -PHEWAS_CODE_HL, -PHEWAS_CODE, -PHEWAS_STRING) %>% 
  distinct()
```

## DISEASE2SYMPTOMS
```{r}
DISEASE2SYMPTOMS  <- tbl(src,"REFERENCE.DISEASE2SYMPTOMS") %>% collect()

# cutoff PUBMED_OCCURRENCE 
DISEASE2SYMPTOMS <- DISEASE2SYMPTOMS %>% filter(PUBMED_OCCURRENCE >= 10) %>% dplyr::select(-PUBMED_OCCURRENCE, -TFIDF_SCORE) %>% distinct()

# MESH_DISEASE_TERM to ICD9
###########################
DISEASE2SYMPTOMS <- as_tibble(apply(DISEASE2SYMPTOMS, 2, function(x) trimws(toupper(x))))
ICD9GROUPING$ICD9_STRING   <- trimws(toupper(ICD9GROUPING$ICD9_STRING))
ICD9GROUPING$PHEWAS_STRING <- trimws(toupper(ICD9GROUPING$PHEWAS_STRING))

# -by ICD9_STRING
# -by PHEWAS_STRING
m1 <- inner_join(DISEASE2SYMPTOMS, ICD9GROUPING %>% dplyr::select(ICD9_STRING, ICD9) %>% distinct(), by= c("MESH_DISEASE_TERM" = "ICD9_STRING"))
m2 <- inner_join(DISEASE2SYMPTOMS, ICD9GROUPING %>% dplyr::select(PHEWAS_STRING, ICD9) %>% distinct(), by= c("MESH_DISEASE_TERM" = "PHEWAS_STRING"))

DISEASE2SYMPTOMS <- rbind(m1,m2) %>%  dplyr::select(-MESH_DISEASE_TERM) %>% distinct()

rm(m1,m2)

```

## DISEASE2GENES
```{r}
load("~/epilepsie_preprocessing_descrStat/output/disease2genes_2016-10-24.RData")
disease2genes <- disease2genes %>% dplyr::rename(score.Disease2Gene = score) %>% distinct()

# gene subelection by gene-disease-score
# (score as least as big as the minimal score of curated gene-disease associations)
min_curated_score <- disease2genes %>% filter(source_descr == "curated") %>% dplyr::select(score.Disease2Gene)
min_curated_score <- min(min_curated_score$score.Disease2Gene)
disease2genes <- disease2genes %>% filter(score.Disease2Gene >= min_curated_score, NofPmids >= 10)

# drop useless columns
disease2genes <- 
  disease2genes %>% 
  dplyr::select(-UMLS, -score.Disease2Gene, -description, - diseaseName, -NofPmids, -PHEWAS_STRING, -source_descr, -geneName) %>%
  distinct()

rm(min_curated_score)

# genes per disease
disease2genes %>% group_by(ICD9) %>% tally(sort = TRUE)
sapply(disease2genes, n_distinct)
```

###  diag.freq to genes
```{r}
# left_join(DISEASES, disease2genes %>% group_by(ICD9) %>% tally(sort = TRUE)) %>% 
#   filter(!is.na(n.y)) %>% 
#   left_join(ICD9GROUPING %>% dplyr::select(ICD9,ICD9_STRING) %>% distinct(), by = "ICD9")
```

# enrichment analysis

## shared code for GO + PATHWAYS
```{r}
# enrichment analysis only for diseases with at least 10 genes
x <- disease2genes %>% group_by(ICD9) %>% tally(sort=TRUE) %>% filter(n>=10)
disease2genes.enr <- disease2genes %>% filter(ICD9 %in% x$ICD9)
diseases.enr <- unique(disease2genes.enr$ICD9)

# diagnosis gene-sets
disease2genes.enr <- 
  sapply(
    diseases.enr,
    FUN = function(disease) disease2genes.enr %>% filter(ICD9 == disease) %>% dplyr::select(entrez)
  )
```

## GO enrichment

### enrichment function
```{r}
library("org.Hs.eg.db")
library("GSEABase")
library("GOstats")

# construct matching table
frame       <- toTable(org.Hs.egGO)
goframeData <-  data.frame(frame$go_id, frame$Evidence, frame$gene_id)
goFrame     <- GOFrame(goframeData, organism <- "Homo sapiens")
goAllFrame  <- GOAllFrame(goFrame) 
gsc         <- GeneSetCollection(goAllFrame, setType = GOCollection()) 

# construct background for hypergemetric testing: all entrez-genes for which GO terms exist in org.Hs.eg.db
keys <- keys(org.Hs.eg.db, keytype = "ENTREZID")
universe <- as_tibble(AnnotationDbi::select(org.Hs.eg.db, keys = keys, keytype="ENTREZID" ,columns=c("ENTREZID","GO")))
universe <- universe[complete.cases(universe),] %>% dplyr::select(ENTREZID) %>% distinct()

# detach("package:org.Hs.eg.db", unload=TRUE)

# define test params, then hypergeom. test, then fdr-correction
GoEnrichment <- function(genes, .universe, .ontology = "BP", .pvalueCutoff = 0.05, .fdrValueCutoff = 0.05) {
  params <- GSEAGOHyperGParams(name="My Custom GSEA based annot Params",
                               geneSetCollection=gsc,
                               geneIds = genes,
                               universeGeneIds = .universe,
                               ontology = .ontology,
                               pvalueCutoff = .pvalueCutoff,
                               conditional = TRUE,
                               testDirection = "over")
  
  enriched.go <- summary(hyperGTest(params))
  fdr         <- fdrtool::fdrtool(x = enriched.go$Pvalue, statistic = "pvalue", plot = FALSE, verbose = FALSE)$qval
  enriched.go <- cbind(enriched.go,fdr) %>% filter( Pvalue <= .pvalueCutoff, fdr <= .fdrValueCutoff)
  
  return(enriched.go)
}
```

### enrichment analysis
```{r}
# enrichment analysis for BP / MF / CC

goEnrBP <- mclapply(disease2genes.enr, FUN = function(genes) GoEnrichment(genes, universe$ENTREZID, "BP"), mc.preschedule = FALSE, mc.cores = 1)

# keep only go-id's with Count > 1 and Size > 1 to avoid over-optimistic enrichment
# keep only go.term column
goEnrBP <- lapply(goEnrBP, function(x) x %>% filter(Count >1, Size >1) %>% dplyr::select(Term) %>% dplyr::rename(GO = Term))

# combine BP + MF + CC + related diseases
l1 <- unlist(lapply(goEnrBP,function(x) dim(x)[1]))
x1 <- data.frame(ICD9 = rep(diseases.enr,l1), unlist(goEnrBP), row.names = NULL, stringsAsFactors = F) %>% setNames(c("ICD9","GO"))
disease2go    <- x1
disease2go    <- NormalizeStrings(disease2go)
disease2go$GO <- make.names(disease2go$GO)
disease2go     <- disease2go %>% distinct()
```

## PATHWAY ENRICHMENT

### enrichment function
```{r}
PathEnrichment <- function(genes, entrez2path, .pvalValueCutoff = 0.2, .fdrValueCutoff = 0.2) {
#’ @param genes vector of entrez-ids for which enrichment analysis should be computed
#’ @param entrez2path dataframe with two columns \code{ENTREZ} and \code{PATH}
#’ @param .fdrValueCutoff cutoff threshold of the false-discovery-rate
  pathsInGenset <- entrez2path %>% filter(ENTREZID %in% genes) %>% dplyr::select(PATH) %>% distinct()
  genesPerPath  <- entrez2path %>% filter(PATH %in% pathsInGenset$PATH) %>% dplyr::select(PATH, ENTREZID) %>% distinct() %>% group_by(PATH) %>% tally() 
  N             <- as.numeric(n_distinct(entrez2path$ENTREZID)) # background for hyperG-Test

if(length(pathsInGenset$PATH) > 0) {
  result <- c()
  for(path in unique(pathsInGenset$PATH)) {
    white.all     <- entrez2path %>% filter(PATH == path) %>% dplyr::select(ENTREZID) %>% distinct() %>% unlist() %>%  as.character()
    size   <- length(white.all)
    count  <- sum(genes %in% white.all)
    pval   <- phyper(count - 1e-3, size, N-size, length(genes), lower.tail = FALSE)
    result <- rbind(result, data.frame(path, count, size, pval))
  }
  fdr    <- fdrtool::fdrtool(x = result$pval, statistic = "pvalue", plot = FALSE, verbose = FALSE)$qval
  result <- cbind(result,fdr) %>% filter( pval <= .pvalValueCutoff, fdr <= .fdrValueCutoff)
  result <- result[order(result$pval),]
  return(result)
}
}
```

### KEGG enrichment
```{r}
keys <- keys(org.Hs.eg.db, keytype = "ENTREZID")
kegg2entrez <- as_tibble(AnnotationDbi::select(org.Hs.eg.db, keys = keys, keytype="ENTREZID" ,columns=c("ENTREZID","PATH")))
kegg2entrez <- kegg2entrez[complete.cases(kegg2entrez),]

disease2path.enr <- mclapply(disease2genes.enr, FUN = function(x) (PathEnrichment(x, kegg2entrez))[,1], mc.preschedule = FALSE, mc.cores = 1)
disease2path.enr <- lapply(disease2path.enr, as.character)

# attached enriched paths to diseases
l <- unlist(lapply(disease2path.enr, length))
disease2path <- data.frame(ICD9 = rep(diseases.enr,l), PATH = unlist(disease2path.enr), stringsAsFactors = F, row.names = NULL)

# id to name
load("~/epilepsie_preprocessing_descrStat/data/KEGG/kegg_id2names.RData")
disease2path <- 
  left_join(disease2path, kegg_id2names, by = c("PATH" = "KEGG.PATH.ID")) %>%
  dplyr::select(-PATH) %>% 
  dplyr::rename(PATH = KEGG.PATH.NAME)

# normalize Strings
disease2path$PATH <- make.names(disease2path$PATH)

# identifier
disease2kegg <- disease2path
```

### WIKI enrichment
```{r}
load("~/epilepsie_preprocessing_descrStat/data/wikipathways/entrez2wiki.RData")

disease2path.enr <- mclapply(disease2genes.enr, FUN = function(x) (PathEnrichment(x, entrez2wiki))[,1], mc.preschedule = FALSE, mc.cores = 1)
disease2path.enr <- lapply(disease2path.enr, as.character)

# attached enriched paths to diseases
l <- unlist(lapply(disease2path.enr, length))
disease2path <- data.frame(ICD9 = rep(diseases.enr,l), PATH = unlist(disease2path.enr), stringsAsFactors = F, row.names = NULL)

# normalize Strings
disease2path$PATH <- make.names(disease2path$PATH)

# identifier
disease2wiki <- disease2path
```

### combine KEGG + WIKI 
```{r}
disease2path <- rbind(disease2kegg,disease2wiki) %>% distinct()
```

# feature matrix

## standardisation: rename columns, prepare tables

* DISEASE2PHEWAS
* DISEASE2PHEWASHL
* DISEASE2SYMPTOMS
* MESH2ICD9
* DISEASE2BIOMARKER
* disease2path (enriched)
* disease2go (enriched)

```{r}
DISEASES          <- DISEASES
DISEASE2PHEWAS    <- DISEASE2PHEWAS %>% dplyr::select(ICD9, PHEWAS_STRING) %>% distinct() %>% dplyr::rename(DISEASE.PHEWAS = PHEWAS_STRING)
DISEASE2PHEWASHL  <- DISEASE2PHEWAS_HL %>% dplyr::select(ICD9, PHEWAS_STRING_HL) %>% distinct() %>% dplyr::rename(DISEASE.PHEWAS_HL = PHEWAS_STRING_HL)
DISEASE2MESH      <- MESH2ICD9 %>% dplyr::rename(DISEASE.MESHDISEASE = mesh, ICD9 = icd9)
DISEASE2SYMPTOMS  <- DISEASE2SYMPTOMS %>% dplyr::rename( DISEASE.SYMPTOMS = MESH_SYMPTOM_TERM)
DISEASE2BIOMARKER <- DISEASE2BIOMARKER %>% dplyr::rename(DISEASE.BIOMARKER = BIOMARKERNAME)
DISEASE2GO        <- disease2go %>% dplyr::rename(DISEASE.GO = GO)
DISEASE2PATH      <- disease2path %>% dplyr::rename(DISEASE.PATH = PATH)

DISEASES$ICD9          <- as.numeric(DISEASES$ICD9)
DISEASE2PHEWAS$ICD9    <- as.numeric(DISEASE2PHEWAS$ICD9)
DISEASE2PHEWASHL$ICD9  <- as.numeric(DISEASE2PHEWASHL$ICD9)
DISEASE2MESH$ICD9      <- as.numeric(DISEASE2MESH$ICD9)
DISEASE2SYMPTOMS$ICD9  <- as.numeric(DISEASE2SYMPTOMS$ICD9)
DISEASE2BIOMARKER$ICD9 <- as.numeric(DISEASE2BIOMARKER$ICD9)
DISEASE2GO$ICD9        <- as.numeric(DISEASE2GO$ICD9)
DISEASE2PATH$ICD9      <- as.numeric(DISEASE2PATH$ICD9)
```
## spread / reshape
```{r}
DISEASE2PHEWAS$val <- 1
DISEASE2PHEWAS <- unique(DISEASE2PHEWAS)
DISEASE2PHEWAS     <- spread(DISEASE2PHEWAS, DISEASE.PHEWAS, val, fill = 0, sep = ".x." )

DISEASE2PHEWASHL$val <- 1
DISEASE2PHEWASHL <- unique(DISEASE2PHEWASHL)
DISEASE2PHEWASHL     <- spread(DISEASE2PHEWASHL, DISEASE.PHEWAS_HL, val, fill = 0, sep = ".x." )

DISEASE2MESH$val <- 1
DISEASE2MESH <- unique(DISEASE2MESH)
DISEASE2MESH     <- spread(DISEASE2MESH, DISEASE.MESHDISEASE, val, fill = 0, sep = ".x." )

DISEASE2SYMPTOMS$val <- 1
DISEASE2SYMPTOMS <- unique(DISEASE2SYMPTOMS)
DISEASE2SYMPTOMS     <- spread(DISEASE2SYMPTOMS, DISEASE.SYMPTOMS, val, fill = 0, sep = ".x." )

DISEASE2BIOMARKER$val <- 1
DISEASE2BIOMARKER <- unique(DISEASE2BIOMARKER)
DISEASE2BIOMARKER     <- spread(DISEASE2BIOMARKER, DISEASE.BIOMARKER, val, fill = 0, sep = ".x." )

DISEASE2GO$val <- 1
DISEASE2GO <- unique(DISEASE2GO)
DISEASE2GO     <- spread(DISEASE2GO, DISEASE.GO, val, fill = 0, sep = ".x." )

DISEASE2PATH$val <- 1
DISEASE2PATH <- unique(DISEASE2PATH)
DISEASE2PATH     <- spread(DISEASE2PATH, DISEASE.PATH, val, fill = 0, sep = ".x." )

# print summary
tables <- c("DISEASE2PHEWAS","DISEASE2PHEWASHL", "DISEASE2MESH", "DISEASE2SYMPTOMS", "DISEASE2BIOMARKER", "DISEASE2GO", "DISEASE2PATH")
dims   <- c(paste(dim(DISEASE2PHEWAS),collapse=" x "),
            paste(dim(DISEASE2PHEWASHL),collapse=" x "),
            paste(dim(DISEASE2MESH),collapse=" x "),
            paste(dim(DISEASE2SYMPTOMS),collapse=" x "),
            paste(dim(DISEASE2BIOMARKER),collapse=" x "),
            paste(dim(DISEASE2GO),collapse=" x "),
            paste(dim(DISEASE2PATH),collapse=" x ")
            )

s           <- as.data.frame(dims,tables)
colnames(s) <- "nICD9 x nFeatures"
knitr::kable(s)
```
## join descriptors to DISEASES
```{r}
DISEASES.FEATMAT <- 
  DISEASES %>% 
  left_join(DISEASE2PHEWAS) %>% 
  left_join(DISEASE2PHEWASHL) %>% 
  left_join(DISEASE2GO) %>% 
  left_join(DISEASE2PATH) %>% 
  left_join(DISEASE2MESH) %>% 
  left_join(DISEASE2SYMPTOMS) %>% 
  left_join(DISEASE2BIOMARKER) 

DISEASES.FEATMAT[is.na(DISEASES.FEATMAT)] <- 0
```

## save
```{r}
save(DISEASES.FEATMAT, file="output/featureMatrix/DISEASE.FEATMAT.RData",compress=TRUE)
```

