---
title: "disease_Descriptors.Rmd"
author: "Thomas Gerlach"
date: "17 11 2016"
output: html_document
---


# load core data and connect to netezza
```{r start, echo=FALSE}
source("z_start.R") # initalize netezza DB, load core data

# if(!exists("TABLE04SAS7BDAT")) load("./data/TABLE04SAS7BDAT.RData")
# PATIENTS_AGE_REGION <- TABLE04SAS7BDAT
```


# script specific functions
```{r}
parseColByIdx <- function(data, colname,idx.key,sep=";") {
  
  colidx          <- which(colnames(data) %in% colname)
  row             <- data[,colidx] != "" | !is.na(data[,colidx])
  y.data          <- data[row,]
  y               <- sapply(y.data[,colidx], function(x) strsplit(x,split = sep))
  y.length        <- sapply(y,length)
  y.data          <- y.data[rep(seq_len(nrow(y.data)), y.length),]
  y.data[,colidx] <- unlist(y)
  
  data     <- data[,-colidx]
  y.data   <- y.data[,c(idx.key,colidx)]
  data.new <- left_join(data,y.data)
  return(unique(data.new))
}
###################
parseMe <- function(x) c(sum(grepl(";",x)), sum(grepl(",",x)))
###################
parseICD9 <- function(icd9.unparsed) {
  
  ICD9GROUPING <- tbl(src,"REFERENCE.ICD9GROUPING") %>% filter(ICD9 != "") %>%   distinct() %>%  collect()
  ICD9GROUPING <- as_tibble(apply(ICD9GROUPING,2,as.character))
  ICD9 <- as_tibble(cbind(ICD9c = ICD9GROUPING$ICD9, ICD9n = ICD9GROUPING$ICD9 %>% as.character() %>% as.numeric()))
  ICD9 <- suppressWarnings(data.frame(ICD9c = ICD9GROUPING$ICD9, ICD9n = as.numeric(as.character(ICD9GROUPING$ICD9))))
  
  parsed <- c()
  
  for(k in 1:length(icd9.unparsed)) { 
    
     if(k %% 100 == 0) print(k)
    #print(k)
    x1 <- strsplit(icd9.unparsed[k],",") %>% unlist() %>% strsplit("-")
    x2 <- unlist(lapply(x1,length))
    
    # parse and attach non-interval elements
    x2.nonint <- which(x2==1)
    if(length(x2.nonint)>0){
      x2.nonint <- trimws(unlist(x1[x2.nonint]))
      parsed    <- rbind(parsed,cbind(icd9.unparsed[k],ICD9=x2.nonint,row.names=NULL))
    }
    
    # parse and attach interval-elements (e.g "120-121")
    x2.interv <- which(x2==2)
    
    if(length(x2.interv) >0) {
      
      for (l in 1:length(x2.interv)) {
        x3 <- unlist(x1[[x2.interv[l]]])
        lb <- as.numeric(x3[1])
        rb <- as.numeric(x3[2])
        if(rb == floor(rb)) rb <- rb + 1 - 1e-3
        x3.ICD9.idx <- which(between(ICD9$ICD9n,lb,rb))
        parsed      <- rbind(parsed,cbind(icd9.unparsed[k],ICD9=as.character(ICD9$ICD9c[x3.ICD9.idx]),row.names=NULL))
      }
    }
  }
  colnames(parsed) <- c("unparsed","parsed")
  return(parsed)
}


```

# markdown defaults
```{r setup, include=FALSE, echo = FALSE }
# set knitr defaults
knitr::opts_chunk$set(echo = FALSE, include = FALSE, fig.width=20, fig.height=10)
```

# load matching resources

## ICD9GROUPING
```{r}
ICD9GROUPING <- tbl(src, "REFERENCE.ICD9GROUPING") %>% filter(ICD9 != "") %>% collect()
```

## MESH2ICD9
```{r, echo = FALSE}
MESH2ICD9 <- read.csv("~/epilepsie_preprocessing_descrStat/data/diseaseDescriptors/doid_MESH2ICD9_2016-09-15.csv", header=TRUE, sep=";")
MESH2ICD9 <- as_tibble(unique(MESH2ICD9))

# drop IDS which no MESH terms
MESH2ICD9$mesh[MESH2ICD9$mesh == ""] <- NA
x <- MESH2ICD9 %>% group_by(id) %>% select(mesh) %>% filter(!is.na(mesh)) %>% distinct()
MESH2ICD9 <- MESH2ICD9 %>% filter(id %in% x$id)

# drop IDs which no ICD9 terms
x <- MESH2ICD9 %>% group_by(id) %>% select(icd9) %>% filter(!is.na(icd9)) %>% distinct()
MESH2ICD9 <- MESH2ICD9 %>% filter(id %in% x$id)

# reshape
MESH2ICD9x <- MESH2ICD9 %>% select(id, mesh) %>% filter(!is.na(mesh))
MESH2ICD9x <- left_join(MESH2ICD9x, MESH2ICD9 %>% select(id, icd9) %>% filter(!is.na(icd9)))
MESH2ICD9  <- MESH2ICD9x %>% select(-id) %>% distinct()

#
rm(MESH2ICD9x, x)
```

## DISEASE2BIOMARKER
```{r}
DISEASE2BIOMARKER <- tbl(src, "REFERENCE.DISEASE2BIOMARKER" ) %>% collect()
DISEASE2BIOMARKER <- DISEASE2BIOMARKER %>% select(-DISEASE, -ICD10, -BIOMARKERNAME) %>% distinct()

# parse ICD9
icd9parsed        <- as_tibble(parseICD9( as.character(unique(DISEASE2BIOMARKER$ICD9)) ))
DISEASE2BIOMARKER <- left_join(DISEASE2BIOMARKER, icd9parsed, by = c("ICD9" = "unparsed")) %>% select(-ICD9) %>% dplyr::rename(ICD9 = parsed) %>% distinct()
#
rm(icd9parsed)
```

## DISEASE2GENES
```{r}
load("~/epilepsie_preprocessing_descrStat/output/disease2genes_2016-10-24.RData")
disease2genes <- disease2genes %>% dplyr::rename(score.Disease2Gene = score) %>% distinct()

# gene subelection by gene-disease-score
# (score as least as big as the minimal score of curated gene-disease associations)
min_curated_score <- disease2genes %>% filter(source_descr == "curated") %>% dplyr::select(score.Disease2Gene)
min_curated_score <- min(min_curated_score$score.Disease2Gene)
disease2genes <- disease2genes %>% filter(score.Disease2Gene >= min_curated_score, NofPmids >= 10)

# drop useless columns
disease2genes <- disease2genes %>% select(-UMLS, -score.Disease2Gene, -description, - diseaseName, -NofPmids, -PHEWAS_STRING, -source_descr) %>% distinct()
#
rm(min_curated_score)
```

## DISEASE2SYMPTOMS

```{r}
DISEASE2SYMPTOMS  <- tbl(src,"REFERENCE.DISEASE2SYMPTOMS") %>% collect()

# cutoff PUBMED_OCCURRENCE 
DISEASE2SYMPTOMS <- DISEASE2SYMPTOMS %>% filter(PUBMED_OCCURRENCE >= 10) %>% select(-PUBMED_OCCURRENCE, -TFIDF_SCORE) %>% distinct()

# MESH_DISEASE_TERM to ICD9
###########################
DISEASE2SYMPTOMS <- as_tibble(apply(DISEASE2SYMPTOMS, 2, function(x) trimws(toupper(x))))
ICD9GROUPING$ICD9_STRING   <- trimws(toupper(ICD9GROUPING$ICD9_STRING))
ICD9GROUPING$PHEWAS_STRING <- trimws(toupper(ICD9GROUPING$PHEWAS_STRING))

# -by ICD9_STRING
# -by PHEWAS_STRING
m1 <- inner_join(DISEASE2SYMPTOMS, ICD9GROUPING %>% select(ICD9_STRING, ICD9) %>% distinct(), by= c("MESH_DISEASE_TERM" = "ICD9_STRING"))
m2 <- inner_join(DISEASE2SYMPTOMS, ICD9GROUPING %>% select(PHEWAS_STRING, ICD9) %>% distinct(), by= c("MESH_DISEASE_TERM" = "PHEWAS_STRING"))

DISEASE2SYMPTOMS <- rbind(m1,m2) %>%  select(-MESH_DISEASE_TERM) %>% distinct()

rm(m1,m2)
```

## ENTREZ2 ...KEGG ...GO
```{r}
library(org.Hs.eg.db)
columns(org.Hs.eg.db)
e                <- keys(org.Hs.eg.db, keytype = "ENTREZID")

# KEGG
ENTREZ2KEGG      <- as_tibble(AnnotationDbi::select(org.Hs.eg.db, keys = e, keytype="ENTREZID" ,columns=c("ENTREZID","PATH")))
ENTREZ2KEGG      <- ENTREZ2KEGG[complete.cases(ENTREZ2KEGG),] %>% distinct()
ENTREZ2KEGG$PATH <- sapply(X = ENTREZ2KEGG$PATH, FUN = function(x) paste("hsa",x,sep=""))

# GO
ENTREZ2GO    <- as_tibble(select(org.Hs.eg.db, keys=e, keytype="ENTREZID" ,columns=c("ENTREZID","GO")))
ENTREZ2GO    <- ENTREZ2GO %>% dplyr::select(-EVIDENCE) %>% distinct()
ENTREZ2GO$GO <- gsub("GO:","",ENTREZ2GO$GO)
ENTREZ2GO    <- ENTREZ2GO[complete.cases(ENTREZ2GO),]
ENTREZ2GO$GO <- paste(ENTREZ2GO$ONTOLOGY,".",ENTREZ2GO$GO,sep="")
ENTREZ2GO    <- ENTREZ2GO %>% dplyr::select(-ONTOLOGY) %>% distinct()

detach("package:org.Hs.eg.db", unload=TRUE)
detach("package:AnnotationDbi", unload=TRUE)

rm(e)
```

# rename columns, prepare tables

* ICD9GROUPING
* MESH2ICD9
* DISEASE2BIOMARKER
* disease2genes
* DISEASE2SYMPTOMS
* ENTREZ2KEGG
* ENTREZ2GO

```{r}
DISEASE2PHEWAS    <- ICD9GROUPING %>% select(ICD9, PHEWAS_CODE) %>%  filter(PHEWAS_CODE != "") %>% distinct()
DISEASE2MESH      <- MESH2ICD9 %>% dplyr::rename(DISEASE.MESHDISEASE = mesh, DISEASE.ICD9 = icd9)
DISEASE2SYMPTOMS  <- DISEASE2SYMPTOMS %>% dplyr::rename(DISEASE.ICD9 = ICD9, DISEASE.MESHSYMPTOM = MESH_SYMPTOM_TERM)
DISEASE2BIOMARKER <- DISEASE2BIOMARKER %>% dplyr::rename(DISEASE.BIOMARKER_REC_ID = BIOMARKER_REC_ID, DISEASE.ICD9 = ICD9)
disease2genes     <- disease2genes %>% dplyr::rename(DISEASE.ENTREZ = entrez, DISEASE.GENESYMBOL = geneName, DISEASE.ICD9 = ICD9)
ENTREZ2KEGG       <- ENTREZ2KEGG %>% dplyr::rename(DISEASE.ENTREZ = ENTREZID, DISEASE.PATH = PATH)
ENTREZ2GO         <- ENTREZ2GO %>% dplyr::rename(DISEASE.ENTREZ = ENTREZID, DISEASE.GO = GO)
```

## DISEASE2PHEWAS, only 1 digit
```{r}
x <- DISEASE2PHEWAS$PHEWAS_CODE %>% as.character() %>% as.numeric() %>% (function(x) floor(10*x)/10)
DISEASE2PHEWAS$PHEWAS_CODE <- x
DISEASE2PHEWAS             <- DISEASE2PHEWAS[complete.cases(DISEASE2PHEWAS),] %>% dplyr::rename(DISEASE.ICD9 = ICD9, DISEASE.PHEWAS = PHEWAS_CODE)

rm(x)
```








## get GENETIC DISEASE DESCRIPTORS

* DISEASE2GENES
* DISEASE2GO: only enriched (hyper-geom. test, FDR)
* DISEASE2KEGG: only enriched (hyper-geom. test, FDR)
```{r}
DIS_GEN_GO_PATH <- 
  left_join(disease2genes, ENTREZ2KEGG) %>% 
  left_join(ENTREZ2GO) %>% 
  select(-DISEASE.ENTREZ) %>% 
  distinct()

# DISEASE2GENES
DISEASE2GENES <- DIS_GEN_GO_PATH %>% select(DISEASE.ICD9, DISEASE.GENESYMBOL) %>% distinct()

# DISEASE2GO
DISEASE2GO <- 
  DIS_GEN_GO_PATH %>% 
  select(DISEASE.ICD9, DISEASE.GENESYMBOL, DISEASE.GO) %>% 
  distinct() %>% 
  select(-DISEASE.GENESYMBOL) %>% 
  group_by(DISEASE.ICD9, DISEASE.GO) %>% 
  tally()

?phyper()

# DISEASE2KEGG
DISEASE2KEGG  <- DIS_GEN_GO_PATH %>% select(DISEASE.ICD9, DISEASE.PATH) %>% distinct()


rm(DIS_GEN_GO_PATH, disease2genes)
```
















# build disease descriptors feature matrix

##  get ICD9 namespace from TABLE04SAS7BDAT
```{r}
DISEASES <- 
  TABLE04SAS7BDAT %>% 
  select(DIAG) %>% 
  filter(DIAG != "", !is.na(DIAG)) %>%
  dplyr::rename(DISEASE.ICD9 = DIAG) %>% 
  distinct()

# parse ICD9 Codes, e.g 0010 -> 001.0
# two leading zeros -> add dot: 00xxx -> 00.xxx
# one leading zero -> add dot: 0xxxx -> 0.xxxxx
###############################################

x00 <- grep("^[0]{2}",DISEASES$DIAG)
x0  <- grep("^[0]{1}[1-9]",DISEASES$DIAG)

DISEASES$DISEASE.ICD9[x00] <- paste("00",substr(DISEASES$DISEASE.ICD9[x00],3,3),".",substr(DISEASES$DISEASE.ICD9[x00],4,10),sep="")
DISEASES$DISEASE.ICD9[x0]  <- paste("0",substr(DISEASES$DISEASE.ICD9[x0],2,3),".",substr(DISEASES$DISEASE.ICD9[x0],4,10),sep="")
DISEASES <- parseICD9(DISEASES)
```

## transform vartype to char
```{r}

# tables list
# DISEASES
# DISEASE2PHEWAS
# DISEASE2GENES
# DISEASE2GO
# DISEASE2KEGG
# DISEASE2MESH
# DISEASE2SYMPTOMS
# DISEASE2BIOMARKER

toChar <- function(x) as_tibble(apply(x,2,function(y) trimws(toupper(y))))

DISEASES          <- toChar(DISEASES)
DISEASE2PHEWAS    <- toChar(DISEASE2PHEWAS)
DISEASE2GENES     <- toChar(DISEASE2GENES)
DISEASE2GO        <- toChar(DISEASE2GO)
DISEASE2KEGG      <- toChar(DISEASE2KEGG)
DISEASE2MESH      <- toChar(DISEASE2MESH)
DISEASE2SYMPTOMS  <- toChar(DISEASE2SYMPTOMS)
DISEASE2BIOMARKER <- toChar(DISEASE2BIOMARKER)

```

# spread / reshape
```{r}


DISEASE2PHEWAS$val <- TRUE
DISEASE2PHEWAS     <- spread(DISEASE2PHEWAS, DISEASE.PHEWAS, val, fill = FALSE, sep = "_x_" )

DISEASE2GENES$val <- TRUE
DISEASE2GENES     <- spread(DISEASE2GENES, DISEASE.GENESYMBOL, val, fill = FALSE, sep = "_x_" )

DISEASE2GO$val <- TRUE
DISEASE2GO     <- spread(DISEASE2GO, DISEASE.GO, val, fill = FALSE, sep = "_x_" )

DISEASE2KEGG$val <- TRUE
DISEASE2KEGG     <- spread(DISEASE2KEGG, DISEASE.PATH, val, fill = FALSE, sep = "_x_" )

DISEASE2MESH$val <- TRUE
DISEASE2MESH     <- spread(DISEASE2MESH, DISEASE.MESHDISEASE, val, fill = FALSE, sep = "_x_" )

DISEASE2SYMPTOMS$val <- TRUE
DISEASE2SYMPTOMS     <- spread(DISEASE2SYMPTOMS, DISEASE.MESHSYMPTOM, val, fill = FALSE, sep = "_x_" )

DISEASE2BIOMARKER$val <- TRUE
DISEASE2BIOMARKER     <- spread(DISEASE2BIOMARKER, DISEASE.BIOMARKER_REC_ID, val, fill = FALSE, sep = "_x_" )

# print summary
tables <- c("DISEASE2PHEWAS", "DISEASE2GENES", "DISEASE2GO", "DISEASE2KEGG", "DISEASE2MESH", "DISEASE2SYMPTOMS", "DISEASE2BIOMARKER")
dims   <- c(paste(dim(DISEASE2PHEWAS),collapse=" x "),
            paste(dim(DISEASE2GENES),collapse=" x "),
            paste(dim(DISEASE2GO),collapse=" x "),
            paste(dim(DISEASE2KEGG),collapse=" x "),
            paste(dim(DISEASE2MESH),collapse=" x "),
            paste(dim(DISEASE2SYMPTOMS),collapse=" x "),
            paste(dim(DISEASE2BIOMARKER),collapse=" x ")
            )
s           <- as.data.frame(dims,tables)
colnames(s) <- "nICD9 x nFeatures"
knitr::kable(s)
```

## join descriptors to DISEASES
```{r}

DISEASES.FEATMAT <- 
  left_join(DISEASES, DISEASE2PHEWAS) %>% 
  left_join(DISEASE2GENES) %>% 
  left_join(DISEASE2GO) %>% 
  left_join(DISEASE2KEGG) %>% 
  left_join(DISEASE2MESH) %>% 
  left_join(DISEASE2SYMPTOMS) %>% 
  left_join(DISEASE2BIOMARKER) 

```

## save
```{r}
save(DISEASES.FEATMAT, file="output/featureMatrix/featMat_diseaseDescr.RData",compress=TRUE)
```

