---
title: "drugRelatedDescriptors"
subtitle: ""
author: "Thomas Gerlach"
date: ""
output: html_document
---

# setup
```{r start, echo=FALSE, include = FALSE}
rm(list=setdiff(ls(), "TABLE04SAS7BDAT"))
knitr::opts_chunk$set(echo = FALSE, fig.width=10, fig.height=7) # markdown
source("z_start.R") # initalize netezza DB, load core data
if(!exists("TABLE04SAS7BDAT")) load("./data/TABLE04SAS7BDAT.RData")
```

# global params
```{r}
params <- vector("list")
params$thres.subst.freq   <- 2
```
# script specific functions
```{r}

####
parseColByIdx <- function(data, colname,sep=";") {
  
  keyIdx  <- which(colnames(data) %in% colname)
  isEmpty <- data[,keyIdx] == "" & is.na(data[,keyIdx])
  data    <- data[!isEmpty,]
  
  parse         <- data[,keyIdx]
  parsed.list   <- sapply(parse, function(x) strsplit(x,split = sep))
  reps          <- sapply(parsed.list,length)
  
  parsed        <- unlist(parsed.list)
  cnames        <- c(colnames(data[,-keyIdx]),colname)
  
  row.reps      <- unlist(mapply(rep, 1:length(reps), reps))
  data.new      <- cbind(data[row.reps, - keyIdx], parsed)
  colnames(data.new) <- cnames
  
  return(data.new)
  
}
###################
parseMe <- function(x) c(sum(grepl(";",x)), sum(grepl(",",x)))
```


# filter patients
```{r standardize population, echo=FALSE}
PATIENTS_AGE_REGION <- FILTER_PATIENTS_AGE_REGION(TABLE04SAS7BDAT)
PATIENTS_AGE_REGION <- PATIENTS_AGE_REGION %>% filter(IXDAYS >= -2*365) %>% distinct()
```

# frequent substances only
```{r start, echo=FALSE}
PATIENTS_AGE_REGION <- FILTER_PATIENTS_AGE_REGION(PATIENTS_AGE_REGION)
subst.freq          <- SUBSTANCE_FREQUENCY(PATIENTS_AGE_REGION)
relevant.subst      <- subst.freq %>% filter(freq >= params$thres.subst.freq) %>% dplyr::select(SUBSTANCENAME)

PATIENTS.DRUGS1 <- 
  PATIENTS_AGE_REGION %>% filter(SUBSTANCENAME %in% relevant.subst$SUBSTANCENAME) %>% 
 dplyr::select(SUBSTANCENAME, PROPRIETARYNAME, PHARMCLS1) %>% 
  distinct()

PATIENTS.DRUGS1 <- as_tibble(apply(PATIENTS.DRUGS1,2,function(x) trimws(toupper(x))))
```

## omit emtpy PHARMCLS1 entries, if drug has at least an non-empty  entry
```{r}
x1 <- PATIENTS.DRUGS1 %>% filter(PHARMCLS1 != "") %>% select(SUBSTANCENAME) %>% distinct()
x2 <- PATIENTS.DRUGS1 %>% filter(SUBSTANCENAME %in% x1$SUBSTANCENAME) %>% select(SUBSTANCENAME, PHARMCLS1) %>% distinct() %>% 
  group_by(SUBSTANCENAME) %>% tally() %>% filter(n>1)
PATIENTS.DRUGS1 <- PATIENTS.DRUGS1 %>% filter( !(PHARMCLS1 == "" & SUBSTANCENAME %in% x2$SUBSTANCENAME) ) %>% distinct()
rm(x1,x2)
```

# add Synonyms
```{r}
DRUG2SYNONYMS <- tbl(src,"REFERENCE.DRUGSYNONYMS") %>% collect() 
DRUG2SYNONYMS <- DRUG2SYNONYMS[DRUG2SYNONYMS$TTDDRUGID != "",]
DRUG2SYNONYMS <- as_tibble(apply(DRUG2SYNONYMS,2, as.character),stringsAsFactors = FALSE)
DRUG2SYNONYMS <- parseColByIdx(DRUG2SYNONYMS,"SYNONYMS")
DRUG2SYNONYMS <- apply(DRUG2SYNONYMS,2, function(x) trimws(toupper(x))) %>% as_tibble() %>% distinct()

PATIENTS.DRUGS2 <-
  left_join(
  PATIENTS.DRUGS1,
  DRUG2SYNONYMS %>% dplyr::select(-TTDDRUGID) %>% distinct(),
  by = c("SUBSTANCENAME" = "DRUGNAME")
  )
```

# SUBSTANCENAME -> TTDDRUGID, SUBSTANCENAME -> names

## PATIENTS.DRUGS2 -> PATIENTS.DRUGS3
```{r}
PATIENTS.DRUGS3 <- 
  rbind(
    PATIENTS.DRUGS2 %>% dplyr::select(SUBSTANCENAME,PHARMCLS1) %>% mutate(names=SUBSTANCENAME),
    PATIENTS.DRUGS2 %>% dplyr::select(SUBSTANCENAME,PHARMCLS1, PROPRIETARYNAME) %>% dplyr::rename(names=PROPRIETARYNAME),
    PATIENTS.DRUGS2 %>% dplyr::select(SUBSTANCENAME,PHARMCLS1, SYNONYMS) %>% dplyr::rename(names=SYNONYMS)
  ) %>% 
  distinct()

PATIENTS.DRUGS3 <- PATIENTS.DRUGS3[complete.cases(PATIENTS.DRUGS3),]
```

## DRUG2SYNONYMS -> DRUG2SYNONYMS2
```{r}
DRUG2SYNONYMS2 <- DRUG2SYNONYMS %>% gather(colnames,names,-TTDDRUGID) %>% dplyr::select(-colnames) %>% distinct()
```

## DRD1
```{r}
DRD1 <- inner_join(PATIENTS.DRUGS3, DRUG2SYNONYMS2)
```

## DRD2
```{r}
DRD2 <- DRD1 %>% dplyr::select(TTDDRUGID, SUBSTANCENAME) %>% distinct()
```

# DRUG -> INDICATION

## DRUG2INDICATIONS -> DRUG2INDICATIONS2
```{r}
DRUG2INDICATIONS2 <- 
  tbl(src,"REFERENCE.DRUG2INDICATIONS") %>% 
  filter(HSP == 1) %>% 
  dplyr::select(DRUG_DESC, ICD9) %>%
  distinct() %>% 
  collect() 

DRUG2INDICATIONS2$DRUG_DESC <- trimws(toupper(DRUG2INDICATIONS$DRUG_DESC))
DRUG2INDICATIONS2$ICD9 <- as.character(DRUG2INDICATIONS$ICD9)

# omit E- and V- ICD9-codings
E <- grep("E",DRUG2INDICATIONS2$ICD9)
V <- grep("V",DRUG2INDICATIONS2$ICD9)
DRUG2INDICATIONS2 <- DRUG2INDICATIONS2[-c(E,V),]
rm(E,V)
```

## DRD2 + DRUG2INDICATIONS2  = drug2indications1
```{r}
drug2indications1 <- inner_join(DRD2, DRUG2INDICATIONS, by=c("SUBSTANCENAME"="DRUG_DESC"))
```

## ICD9GROUPING -> ICD9GROUPING2
### PHEWAS_CODE -> PHEWAS_HIGHLEVEL_CODE
### PHEWAS_HIGHLEVEL_CODE -> PHEWAS_HIGHLEVEL_STRING

```{r}
ICD9GROUPING <- tbl(src,"REFERENCE.ICD9GROUPING") %>% select(ICD9, PHEWAS_CODE, PHEWAS_STRING) %>% distinct() %>% collect()
ICD9GROUPING <- sapply(ICD9GROUPING, function(x) trimws(toupper(as.character(x)))) %>% as_tibble()
ICD9GROUPING$ICD9 <- as.character(ICD9GROUPING$ICD9)
ICD9GROUPING$PHEWAS_STRING <- as.character(ICD9GROUPING$PHEWAS_STRING)
ICD9GROUPING$PHEWAS_STRING <- trimws(toupper(ICD9GROUPING$PHEWAS_STRING))

# omit E- and V- ICD9-codings
E <- grep("E",ICD9GROUPING$ICD9)
V <- grep("V",ICD9GROUPING$ICD9)
ICD9GROUPING <- ICD9GROUPING[-c(E,V),]
rm(E,V)

# erase / unclear missing values
ICD9GROUPING <- ICD9GROUPING %>% filter(ICD9 != "", !is.na(PHEWAS_CODE), PHEWAS_STRING != "", PHEWAS_STRING != "1", PHEWAS_STRING != "-")

# PHEWAS_CODE -> PHEWAS_HIGHLEVEL_CODE
# PHEWAS_HIGHLEVEL_CODE -> PHEWAS_HIGHLEVEL_STRING
ICD9GROUPING$PHEWAS_CODE <- as.numeric(ICD9GROUPING$PHEWAS_CODE)
PHEWAS_HL <- 
  ICD9GROUPING %>% 
  mutate(PHEWAS_floor = floor(PHEWAS_CODE)) %>% 
  mutate(isHighlevel = PHEWAS_floor == PHEWAS_CODE) %>% 
  filter(isHighlevel == TRUE) %>% 
  select(PHEWAS_CODE, PHEWAS_STRING) %>%
  distinct() %>% 
  rename(PHEWAS_CODE_HL = PHEWAS_CODE, PHEWAS_STRING_HL = PHEWAS_STRING)

PHEWAS_HL <- 
  ICD9GROUPING %>% 
  mutate(PHEWAS_CODE_HL = floor(PHEWAS_CODE)) 

# ICD9GROUPING2
ICD9GROUPING2 <- inner_join(ICD9GROUPING2, PHEWAS_HL)
```

# drug2indications1 + ICD9GROUPING2 = drug2indications
```{r}
drug2indications <-  inner_join(ICD9GROUPING2, drug2indications1, by="ICD9") %>% select(-ICD9) %>% distinct()
```





## DRUGSYNONYMSDRUGBANK
```{r}
parseMe(DRUG2SYNONYMS$DRUGNAME)
DRUG2SYNONYMS <- parseColByIdx(DRUG2SYNONYMS,"DRUGNAME")
```
## PATIENTS.DRUGS1 + DRUGSYNONYMSDRUGBANK + DRUG2SYNONYMS
```{r}

```



## DRUG2INDICATIONS, parsing, preprocessing
```{r}
DRUG2INDICATIONS <- tbl(src,"REFERENCE.DRUG2INDICATIONS") %>% collect() 
glimpse(DRUG2INDICATIONS)
DRUG2INDICATIONS <- DRUG2INDICATIONS %>% dplyr::select(DRUG_DESC, ICD9) %>% distinct()
parseMe(DRUG2INDICATIONS$DRUG_DESC)
parseMe(DRUG2INDICATIONS$ICD9)
DRUG2INDICATIONS$DRUG_DESC <- trimws(toupper(DRUG2INDICATIONS$DRUG_DESC))
```

## DRUG2INDICATIONSTTD, parsing , preprocessing
```{r}
D2I <- tbl(src,"REFERENCE.DRUG2INDICATIONSTTD") %>% filter(TTDDRUGID != "", LNM != "") %>%  
  dplyr::select(TTDDRUGID, LNM, ICD9) %>% distinct() %>%  collect()
bugged.rows <- c("280-285, 585.9585.1-585.5403",
                 "272.6, 584, 585, 585.9585.1-585.5403",
                 "584, 585, 585.9585.1-585.5403")
bugged.rows <- mapply(grep, pattern=bugged.rows, MoreArgs=list(x=D2I$ICD9))
bugged.rows <- as.numeric(unlist(bugged.rows))

D2I           <- D2I[-bugged.rows,]
D2I           <- apply(D2I,2, function(x) toupper(trimws(x))) 
D2I           <- as_tibble(D2I)
glimpse(D2I)
D2I$ICD9[D2I$ICD9=="."] <- NA
D2I <- D2I[complete.cases(D2I),]

# parse LNM
###########
parseMe(D2I$LNM)

# parse ICD9
############
D2I <- ParseICD9(D2I %>% dplyr::rename(DIAG = ICD9)) %>% dplyr::rename(ICD9 = DIAG)

# match parsed ICD9 to D2I
D2I <- left_join(D2I, as.data.frame(icd9.parsed,stringsAsFactors=FALSE),by=c("ICD9"="V1"))
D2I <- D2I %>% dplyr::select(-ICD9) %>% dplyr::rename(ICD9 = ICD9.y) %>% distinct()

```

## DRUG2INDICATIONSTTD <-> DRUG2SYNONYMS and DRUG2INDICATIONS <-> DRUG2SYNONYMS
```{r}

# DRUG2INDICATIONSTTD <-> DRUG2SYNONYMS
DRD.A.1   <- left_join(DRUG2SYNONYMS, D2I)
DRD.A.1   <- 
  rbind(
    DRD.A.1 %>% dplyr::select(TTDDRUGID, SYNONYMS, ICD9) %>%  dplyr::rename(DRUGNAME = SYNONYMS) %>% distinct(),
    DRD.A.1 %>% dplyr::select(TTDDRUGID, DRUGNAME, ICD9) %>%  distinct(),
    DRD.A.1 %>% dplyr::select(TTDDRUGID, LNM, ICD9)      %>%  dplyr::rename(DRUGNAME = LNM) %>% distinct()
  ) %>% 
  distinct()

DRD.A.2 <- left_join( DRD.A.1 %>% dplyr::select(-ICD9) %>% distinct(), DRUG2INDICATIONS, by=c("DRUGNAME"="DRUG_DESC"))
DRD.A.2 <- DRD.A.2 %>% filter(!is.na(ICD9))
DRD.A.2 <- rbind(DRD.A.1, DRD.A.2) %>% distinct()

# DRUG2INDICATIONS <-> DRUG2SYNONYMS
DRD.A.3.1 <- left_join(DRUG2SYNONYMS, DRUG2INDICATIONS, by = c("SYNONYMS"="DRUG_DESC")) %>% filter(!is.na(ICD9))
DRD.A.3.2 <- left_join(DRUG2SYNONYMS, DRUG2INDICATIONS, by = c("DRUGNAME"="DRUG_DESC")) %>% filter(!is.na(ICD9))

DRD.A.3.1 <- 
  rbind(
    DRD.A.3.1 %>% dplyr::select(-SYNONYMS) %>% distinct(),
    DRD.A.3.1 %>% dplyr::select(-DRUGNAME) %>% dplyr::rename(DRUGNAME = SYNONYMS) %>% distinct()
  ) %>% distinct()

DRD.A.3.2 <- 
  rbind(
    DRD.A.3.2 %>% dplyr::select(-SYNONYMS) %>% distinct(),
    DRD.A.3.2 %>% dplyr::select(-DRUGNAME) %>% dplyr::rename(DRUGNAME = SYNONYMS) %>% distinct()
  ) %>% distinct()

DRD.A.3 <- rbind(DRD.A.3.1, DRD.A.3.2) %>% distinct()

#
DRD.A.4 <- rbind(DRD.A.2, DRD.A.3) %>% distinct()

# tidy up
rm(DRD.A.2, DRD.A.3.1, DRD.A.3.2, DRD.A.3 )

```

# WAY 0

## PATIENTS_AGE_REGION, capture drugfeatures
```{r}
PATIENTS.DRUGS <-
  PATIENTS_AGE_REGION  %>%
  filter(SUBSTANCENAME != "") %>% 
  dplyr::select(SUBSTANCENAME, PROPRIETARYNAME, PHARMCLS1) %>% 
  distinct()
glimpse(PATIENTS.DRUGS)

PATIENTS.DRUGS <- as_tibble(apply(PATIENTS.DRUGS,2,function(x) trimws(toupper(x))))
```


## DRUGSYNONYMSDRUGBANK, parsing, preprocessing
```{r}
# parse column Synonyms
DRUGSYNONYMSDRUGBANK <- read.delim(file="data/DrugBank/DrugSynonymsDrugBank.csv",sep = ",",header = TRUE,stringsAsFactors = FALSE)
glimpse(DRUGSYNONYMSDRUGBANK)
DRUGSYNONYMSDRUGBANK <- DRUGSYNONYMSDRUGBANK %>% dplyr::select(DrugBank.ID, Common.name,Synonyms) %>% distinct()
DRUGSYNONYMSDRUGBANK <- parseColByIdx(as_tibble(DRUGSYNONYMSDRUGBANK),colname = "Synonyms",sep = "\\|")
glimpse(DRUGSYNONYMSDRUGBANK)
DRUGSYNONYMSDRUGBANK <- DRUGSYNONYMSDRUGBANK[!is.na(DRUGSYNONYMSDRUGBANK$Synonyms),]

# string standardisation
DRUGSYNONYMSDRUGBANK <- as_tibble(apply(DRUGSYNONYMSDRUGBANK,2, function(x) trimws(toupper(x)))) %>% distinct()
DRUGSYNONYMSDRUGBANK <- parseColByIdx(DRUGSYNONYMSDRUGBANK,colname = "Synonyms")
```

## DRUGSYNONYMSDRUGBANK <-> PATIENTS.DRUGS -> PATIENTS.DRUGS
```{r}

PATIENTS.DRUGS.1 <- left_join(PATIENTS.DRUGS, DRUGSYNONYMSDRUGBANK, by=c("SUBSTANCENAME"="Common.name"))
PATIENTS.DRUGS.2 <- left_join(PATIENTS.DRUGS, DRUGSYNONYMSDRUGBANK, by=c("PROPRIETARYNAME"="Common.name"))
PATIENTS.DRUGS <- rbind(PATIENTS.DRUGS.1, PATIENTS.DRUGS.2) %>% distinct()
rm(PATIENTS.DRUGS.1, PATIENTS.DRUGS.2)

```

## PATIENTS.DRUGS <-> DRD.x -> DRD.x+1
```{r}
n0 <- PATIENTS.DRUGS %>% dplyr::select(SUBSTANCENAME, PHARMCLS1)
n1 <- cbind(n0, PATIENTS.DRUGS %>% dplyr::select(SUBSTANCENAME) %>% dplyr::rename(DRUGNAME = SUBSTANCENAME))
n2 <- cbind(n0, PATIENTS.DRUGS %>% dplyr::select(PROPRIETARYNAME) %>% dplyr::rename(DRUGNAME = PROPRIETARYNAME))
n3 <- cbind(n0, PATIENTS.DRUGS %>% dplyr::select(Synonyms) %>% dplyr::rename(DRUGNAME = Synonyms))
PATIENTS.DRUGS <- as_tibble(unique(rbind(n1,n2,n3)))
PATIENTS.DRUGS <- PATIENTS.DRUGS[!is.na(PATIENTS.DRUGS$DRUGNAME),]
rm(n0,n1,n2,n3)

DRD.W0.1 <- inner_join(PATIENTS.DRUGS, DRD.A.4) %>% dplyr::select(-DRUGNAME) %>% distinct()

# add drugname from DRUG2SYNONYMS
DRD.W0.1 <- left_join(DRD.W0.1, DRUG2SYNONYMS %>% dplyr::select(-SYNONYMS)) %>% distinct()
```

## add PHEWAS-Code
```{r}
ICD9GROUPING <- tbl(src,"REFERENCE.ICD9GROUPING") %>% filter(ICD9 != "", PHEWAS_CODE != "") %>% 
  dplyr::select(ICD9, PHEWAS_CODE) %>% distinct() %>% collect()
DRD.W0.2 <- left_join(DRD.W0.1, ICD9GROUPING)
n_distinct(DRD.W0.2$ICD9)
n_distinct(DRD.W0.2$PHEWAS_CODE)

```

# WAY1

## DISEASE2GENES, parsing, preprocessing
```{r}
load("~/epilepsie_preprocessing_descrStat/output/disease2genes_2016-10-24.RData")
disease2genes <- disease2genes %>% dplyr::rename(score.Disease2Gene = score) %>% distinct()
glimpse(disease2genes)

# gene subelection by gene-disease-score
# (score as least as big as the minimal score of curated gene-disease associations)
min_curated_score <- disease2genes %>% filter(source_descr == "curated") %>% dplyr::select(score.Disease2Gene)
min_curated_score <- min(min_curated_score$score.Disease2Gene)
disease2genes <- disease2genes %>% filter(score.Disease2Gene >= min_curated_score, NofPmids >= 10)

# drop unessential columns
disease2genes <- disease2genes %>% dplyr::select(entrez, ICD9) %>% distinct()
```

## DISEASE2GENES
```{r}
# attach to DRD
DRD.W1.1 <- left_join(DRD.W0.2, disease2genes)
```

## DRD, DRUG2SIDEEFFECTS
```{r}
# DRD.W1.1$UMLS <- trimws(toupper(DRD.W1.1$UMLS))
# DRD.W1.2 <- left_join(DRD.W1.1, DRUG2SIDEEFFECTS, by="UMLS")
# glimpse(DRD.W1.2)
# DRD.W1.2 <- DRD.W1.2 %>% filter(!is.na(UMLS)) %>% distinct()

```

## add GO Terms
```{r}
# get matching table
####################

library(org.Hs.eg.db)

e            <- unique(DRD.W1.1$entrez[!is.na(DRD.W1.1$entrez)])
entrez2go    <- select(org.Hs.eg.db, keys=e, keytype="ENTREZID" ,columns=c("ENTREZID","GO"))
entrez2go    <- entrez2go %>% dplyr::select(ENTREZID, GO, ONTOLOGY) %>% distinct()
entrez2go$GO <- gsub("GO:","",entrez2go$GO)

# match DRD.W1.2 <-> entrez2go
##############################
DRD.W1.3 <- left_join(
  DRD.W1.1 %>%  distinct(),
  entrez2go,
  by=c("entrez"="ENTREZID"))

# reformat GO column
# DRD.W1.3    <- DRD.W1.3 %>% dplyr::select(-entrez) %>% distinct()
x              <- !is.na(DRD.W1.3$GO)
DRD.W1.3$GO[x] <- paste(DRD.W1.3$ONTOLOGY[x],".",DRD.W1.3$GO[x],sep="")
DRD.W1.3       <- DRD.W1.3 %>% dplyr::select(-ONTOLOGY)

# add kegg pathways
###################
library(help=org.Hs.eg.db)

keys         <- unique(DRD.W1.3$entrez[!is.na(DRD.W1.3$entrez)])
mapped_genes <- mappedkeys(org.Hs.egPATH)
mapped_genes <- mapped_genes[mapped_genes %in% keys]
kegg         <- as.list(org.Hs.egPATH[mapped_genes])
kegg.l       <- sapply(kegg,length)
entrez       <- mapply(rep,mapped_genes,kegg.l)
entrez2kegg  <- as_tibble(cbind(entrez=unlist(entrez),kegg_id = unlist(kegg)))

DRD.W1.3 <- left_join(DRD.W1.3, entrez2kegg)
DRD.W1.3 <- DRD.W1.3 %>% dplyr::select(TTDDRUGID, entrez, GO, kegg_id) %>% distinct()

# postprocessing
DRD.W1.3 <- DRD.W1.3 %>% dplyr::rename(INDC.entrez = entrez, INDC.GO = GO, INDC.PATH = kegg_id)
#DRD.W1.3$INDC.PATH[!is.na(DRD.W1.3$INDC.PATH)]  <- sapply(DRD.W1.3$INDC.PATH[!is.na(DRD.W1.3$INDC.PATH)], function(x) paste("KEGG.",x,collapse=""))
library(parallel)
DRD.W1.3$INDC.PATH[!is.na(DRD.W1.3$INDC.PATH)] <- 
  unlist(
    mclapply(
      DRD.W1.3$INDC.PATH[!is.na(DRD.W1.3$INDC.PATH)],
      function(x) paste("KEGG.",x,sep="",collapse=""),
      mc.cores = 10)
    )

```



# WAY2, SIDEEFFECTS

## TARGET_CROSSMATCHING, parse SuperDrugATC, DrugName
```{r}
load("~/epilepsie_preprocessing_descrStat/data/TTD/TARGET_CROSSMATCHING.RData")
TARGET_CROSSMATCHING <- TARGET_CROSSMATCHING %>%  filter(SuperDrugATC != "") %>% 
  dplyr::select(TTDID, DrugName, SuperDrugATC) %>% distinct()
sum(grepl(";",TARGET_CROSSMATCHING$DrugName))
sum(grepl(";",TARGET_CROSSMATCHING$SuperDrugATC))
n_distinct(TARGET_CROSSMATCHING$SuperDrugATC)
glimpse(TARGET_CROSSMATCHING)

# parse DrugName
parseMe(TARGET_CROSSMATCHING$DrugName)
TARGET_CROSSMATCHING <- parseColByIdx(TARGET_CROSSMATCHING,"DrugName",sep = ";")

# parse SuperDrugATC
parseMe(TARGET_CROSSMATCHING$SuperDrugATC)
TARGET_CROSSMATCHING <- parseColByIdx(TARGET_CROSSMATCHING,"SuperDrugATC",sep = ";")

```

## DRUGIDS, preprocessing
```{r}
DRUGIDS <- tbl(src,"REFERENCE.DRUGIDS") %>% filter(SOURCE_NAME == "ATC") %>% dplyr::select(-SOURCE_NAME) %>% distinct() %>% collect() %>% 
  dplyr::rename(SuperDrugATC = SOURCE_ID)

parseMe(DRUGIDS$SuperDrugATC)

```


## DRUG2SIDEEFFECTS, parse and transform FREQUENCY
```{r}
DRUG2SIDEEFFECTS <- tbl(src,"REFERENCE.DRUG2SIDEEFFECTS") %>% collect()
DRUG2SIDEEFFECTS <- DRUG2SIDEEFFECTS %>% dplyr::select(STITCHID_FLAT, STITCHID_STEREO, FREQUENCY, MEDDRA, -UMLS) %>% distinct()
DRUG2SIDEEFFECTS <- as_tibble(apply(DRUG2SIDEEFFECTS,2,as.character))
DRUG2SIDEEFFECTS$FREQUENCY <- gsub("%","",DRUG2SIDEEFFECTS$FREQUENCY)
DRUG2SIDEEFFECTS$FREQUENCYn <- as.numeric(DRUG2SIDEEFFECTS$FREQUENCY)

# drop useless rows
x <- which(DRUG2SIDEEFFECTS$FREQUENCY == "postmarketing")
DRUG2SIDEEFFECTS <- DRUG2SIDEEFFECTS[-x,]

# split source
DRUG2S.numeric      <- DRUG2SIDEEFFECTS[!is.na(DRUG2SIDEEFFECTS$FREQUENCYn),]
DRUG2S.nonnumeric   <- DRUG2SIDEEFFECTS[is.na(DRUG2SIDEEFFECTS$FREQUENCYn),]

# # non-numeric: replace unessential
DRUG2S.nonnumeric$FREQUENCY <- gsub("<","",DRUG2S.nonnumeric$FREQUENCY)

# non-numeric: substitute ranges through mean, sep="-"
subst <- grep("-",DRUG2S.nonnumeric$FREQUENCY)
subst.split <- strsplit(DRUG2S.nonnumeric$FREQUENCY[subst],"-")
subst.mean <- unlist(lapply(subst.split, FUN = function(x) mean(as.numeric(x),na.rm=TRUE)))
DRUG2S.nonnumeric$FREQUENCY[subst] <- subst.mean

# non-numeric: substitute ranges through mean, sep="to"
subst <- grep("to",DRUG2S.nonnumeric$FREQUENCY)
subst.split <- strsplit(DRUG2S.nonnumeric$FREQUENCY[subst],"to")
subst.mean <- unlist(lapply(subst.split, FUN = function(x) mean(as.numeric(x),na.rm=TRUE)))
DRUG2S.nonnumeric$FREQUENCY[subst] <- subst.mean

# substitute WHO wording through mean frequency
subst.very_rare <- 0.05
subst.rare <- mean(0.1,0.01)
subst.uncommon <- mean(0.1,1)
subst.infrequent <- subst.uncommon
subst.common <- mean(1,10)
subst.frequent <- subst.common
subst.verycommon <- 10
subst.veryfrequent <- subst.verycommon
subst.01feb <- mean(1,2)
subst.aug66 <- mean(8,66)
subst.okt13 <- mean(10,13)
subst.01jun <- mean(1,6)
subst.jan13 <- mean(1,13)

DRUG2S.nonnumeric$FREQUENCY[DRUG2S.nonnumeric$FREQUENCY == "very rare"] <- subst.very_rare
DRUG2S.nonnumeric$FREQUENCY[DRUG2S.nonnumeric$FREQUENCY == "rare"] <- subst.rare
DRUG2S.nonnumeric$FREQUENCY[DRUG2S.nonnumeric$FREQUENCY == "uncommon"] <- subst.uncommon
DRUG2S.nonnumeric$FREQUENCY[DRUG2S.nonnumeric$FREQUENCY == "infrequent"] <- subst.infrequent
DRUG2S.nonnumeric$FREQUENCY[DRUG2S.nonnumeric$FREQUENCY == "common"] <- subst.common
DRUG2S.nonnumeric$FREQUENCY[DRUG2S.nonnumeric$FREQUENCY == "frequent"] <- subst.common
DRUG2S.nonnumeric$FREQUENCY[DRUG2S.nonnumeric$FREQUENCY == "very common"] <- subst.verycommon
DRUG2S.nonnumeric$FREQUENCY[DRUG2S.nonnumeric$FREQUENCY == "very frequent"] <- subst.veryfrequent
DRUG2S.nonnumeric$FREQUENCY[DRUG2S.nonnumeric$FREQUENCY == "01. Feb"] <- subst.01feb
DRUG2S.nonnumeric$FREQUENCY[DRUG2S.nonnumeric$FREQUENCY == "Aug 66"] <- subst.aug66
DRUG2S.nonnumeric$FREQUENCY[DRUG2S.nonnumeric$FREQUENCY == "Okt 13"] <- subst.okt13
DRUG2S.nonnumeric$FREQUENCY[DRUG2S.nonnumeric$FREQUENCY == "01. Jun"] <- subst.01jun
DRUG2S.nonnumeric$FREQUENCY[DRUG2S.nonnumeric$FREQUENCY == "Jan 13"] <- subst.jan13

DRUG2S.nonnumeric$FREQUENCYn <- DRUG2S.nonnumeric$FREQUENCY

# merge split
DRUG2SIDEEFFECTS <- rbind(DRUG2S.nonnumeric %>% dplyr::select(-FREQUENCY), DRUG2S.numeric %>% dplyr::select(-FREQUENCY))
DRUG2SIDEEFFECTS$FREQUENCYn <- as.numeric(DRUG2SIDEEFFECTS$FREQUENCYn)

# aggregate by mean frequency
DRUG2SIDEEFFECTS <- aggregate(FREQUENCYn ~., data = DRUG2SIDEEFFECTS, FUN=mean, na.rm=TRUE)

# # transform FREQUENCYn to WHO classification
# ############################################
# 
# #  # WHO frequency classification
# # freq  <- c("very common", "common", "uncommon", "rare", "very rare")
# # freq.p <- c(">= 10%", ">= 1% and <10%", ">= 0.1% and < 1%", ">= 0.01% and < 0.1%", "< 0.01%")
# # WHO <- as.data.frame(cbind(freq,freq.p), stringsAsFactors = FALSE)
# 
# very_rare   <- DRUG2SIDEEFFECTS$FREQUENCYn < 0.01
# rare        <- DRUG2SIDEEFFECTS$FREQUENCYn >= 0.01 & DRUG2SIDEEFFECTS$FREQUENCYn  < 0.1
# uncommon    <- DRUG2SIDEEFFECTS$FREQUENCYn >= 0.1  & DRUG2SIDEEFFECTS$FREQUENCYn  < 1
# common      <- DRUG2SIDEEFFECTS$FREQUENCYn >= 1    & DRUG2SIDEEFFECTS$FREQUENCYn  < 10
# very_common <- DRUG2SIDEEFFECTS$FREQUENCYn >= 10
# 
# DRUG2SIDEEFFECTS$WHO <- NA
# DRUG2SIDEEFFECTS$WHO[very_rare]   <- "very_rare"
# DRUG2SIDEEFFECTS$WHO[rare]        <- "rare"
# DRUG2SIDEEFFECTS$WHO[uncommon]    <- "uncommon"
# DRUG2SIDEEFFECTS$WHO[common]      <- "common"
# DRUG2SIDEEFFECTS$WHO[very_common] <- "very_common"
# 
# # standardise strings
DRUG2SIDEEFFECTS$MEDDRA <- trimws(toupper(DRUG2SIDEEFFECTS$MEDDRA))
# DRUG2SIDEEFFECTS$WHO    <- trimws(toupper(DRUG2SIDEEFFECTS$WHO))
# DRUG2SIDEEFFECTS        <- as_tibble(DRUG2SIDEEFFECTS)
# head(DRUG2SIDEEFFECTS)
# 
# # tidy up
# rm(DRUG2S.numeric, DRUG2S.nonnumeric, very_rare, rare, uncommon, common, very_common)
# rm(list=ls(pattern="subst\\."))
# rm(x, subst)
```

## multi-join DRD <-> TARGET_CROSSMATCHING <-> DRUGIDS <-> DRUG2SIDEEFFECTS --> DRD
```{r}

# DRD <-> TARGET_CROSSMATCHING
DRD.W2.1 <- DRD.W0.2 %>% dplyr::select(TTDDRUGID, DRUGNAME) %>% distinct()
DRD.W2.2 <- left_join(DRD.W2.1, TARGET_CROSSMATCHING, by=c("TTDDRUGID"="TTDID"))

DRD.W2.1$DRUGNAME             <- trimws(toupper(DRD.W2.1$DRUGNAME))
TARGET_CROSSMATCHING$DrugName <- trimws(toupper(TARGET_CROSSMATCHING$DrugName))
DRD.W2.3                      <- left_join(DRD.W2.1, TARGET_CROSSMATCHING, by=c("DRUGNAME"="DrugName"))
glimpse(DRD.W2.2)
glimpse(DRD.W2.3)

DRD.W2.4 <- 
  unique(
    rbind(
      DRD.W2.2 %>% dplyr::select(TTDDRUGID, DRUGNAME, SuperDrugATC) %>% distinct(),
      DRD.W2.3 %>% dplyr::select(TTDDRUGID, DRUGNAME, SuperDrugATC) %>% distinct()
      ))

DRD.W2.4 <- DRD.W2.4 %>% filter(!is.na(SuperDrugATC))

# <-> DRUGIDs
DRD.W2.4$SuperDrugATC <- trimws(toupper(DRD.W2.4$SuperDrugATC))
DRUGIDS$SuperDrugATC  <- trimws(toupper(DRUGIDS$SuperDrugATC))
DRD.W2.5              <- left_join(DRD.W2.4, DRUGIDS)

# <-> DRUG2SIDEEFFECTS
DRD.W2.5$FLAT_CHEMICAL   <- trimws(toupper(DRD.W2.5$FLAT_CHEMICAL))
DRD.W2.5$STEREO_CHEMICAL <- trimws(toupper(DRD.W2.5$STEREO_CHEMICAL))
DRUG2SIDEEFFECTS$STITCHID_FLAT <- trimws(toupper(DRUG2SIDEEFFECTS$STITCHID_FLAT))
DRUG2SIDEEFFECTS$STITCHID_STEREO <- trimws(toupper(DRUG2SIDEEFFECTS$STITCHID_STEREO))
DRD.W2.6.1 <- left_join(DRD.W2.5, DRUG2SIDEEFFECTS %>% dplyr::select(-STITCHID_STEREO), by=c("FLAT_CHEMICAL"="STITCHID_FLAT"))
DRD.W2.6.2 <- left_join(DRD.W2.5, DRUG2SIDEEFFECTS %>% dplyr::select(-STITCHID_FLAT), by=c("STEREO_CHEMICAL"="STITCHID_STEREO"))

# combine
DRD.W2.6   <-
  unique(rbind(DRD.W2.6.1, DRD.W2.6.2)) %>% dplyr::select(-FLAT_CHEMICAL,
  -STEREO_CHEMICAL,
  -DRUGNAME,
  -SuperDrugATC
  ) %>%  distinct()

x         <- which(is.na(DRD.W2.6$MEDDRA) & is.na(DRD.W2.6$FREQUENCYn) )
DRD.W2.6  <- DRD.W2.6[-x,] %>% distinct()

# transform FREQUENCYn to WHO classification
############################################

#  # WHO frequency classification
# freq  <- c("very common", "common", "uncommon", "rare", "very rare")
# freq.p <- c(">= 10%", ">= 1% and <10%", ">= 0.1% and < 1%", ">= 0.01% and < 0.1%", "< 0.01%")
# WHO <- as.data.frame(cbind(freq,freq.p), stringsAsFactors = FALSE)

# aggregate
DRD.W2.6 <- as_tibble(aggregate(FREQUENCYn ~., data = DRD.W2.6, FUN = mean))

# transform
very_rare   <- DRD.W2.6$FREQUENCYn < 0.01
rare        <- DRD.W2.6$FREQUENCYn >= 0.01 & DRD.W2.6$FREQUENCYn  < 0.1
uncommon    <- DRD.W2.6$FREQUENCYn >= 0.1  & DRD.W2.6$FREQUENCYn  < 1
common      <- DRD.W2.6$FREQUENCYn >= 1    & DRD.W2.6$FREQUENCYn  < 10
very_common <- DRD.W2.6$FREQUENCYn >= 10

DRD.W2.6$WHO <- NA
DRD.W2.6$WHO[very_rare]   <- 0 # "very_rare"
DRD.W2.6$WHO[rare]        <- 1 # "rare"
DRD.W2.6$WHO[uncommon]    <- 2 # "uncommon"
DRD.W2.6$WHO[common]      <- 3 # "common"
DRD.W2.6$WHO[very_common] <- 4 # "very_common"

# postprocessing
################
DRD.W2.6 <- DRD.W2.6 %>% dplyr::select(-FREQUENCYn) %>% distinct()

```


# WAY 3

## expand drug namespace
```{r}

DRD.W3.1 <-
  rbind(
    DRD.W0.2 %>% dplyr::select(TTDDRUGID, SUBSTANCENAME) %>% distinct() %>% dplyr::rename(NAME = SUBSTANCENAME),
    DRD.W0.2 %>% dplyr::select(TTDDRUGID, DRUGNAME) %>% distinct() %>% dplyr::rename(NAME = DRUGNAME)
  ) %>% distinct()

DRD.W3.1.x <- left_join(DRD.W3.1, DRUGSYNONYMSDRUGBANK, by = c("NAME"="Common.name"))
DRD.W3.1  <- 
  rbind(
    DRD.W3.1.x %>% dplyr::select(-Synonyms),
    DRD.W3.1.x %>% dplyr::select(-NAME) %>% dplyr::rename(NAME = Synonyms)
  ) %>% distinct()

# add Uniprot.ID via DRUGTARGETSDRUGBANK
DRUGTARGETSDRUGBANK <- read.csv("~/epilepsie_preprocessing_descrStat/data/DrugBank/DrugTargetsDrugBank.csv", stringsAsFactors=FALSE)
head(DRUGTARGETSDRUGBANK)
DRUGTARGETSDRUGBANK <- DRUGTARGETSDRUGBANK %>% dplyr::select(DrugBank.ID, UniProt.ID) %>% distinct()

DRD.W3.1e <- left_join(DRD.W3.1, DRUGTARGETSDRUGBANK) %>% dplyr::select(-DrugBank.ID) %>% distinct()
```

## TARGET ONTOLOGY
```{r}

# TARGET ONTOLOGY, load and preprocess
load("~/epilepsie_preprocessing_descrStat/data/TTD/TARGET_ONTOLOGY.RData")
TARGET_ONTOLOGY <- TARGET_ONTOLOGY[,-10]
TARGET_ONTOLOGY <- as_tibble(TARGET_ONTOLOGY)
glimpse(TARGET_ONTOLOGY)
TARGET_ONTOLOGY <- TARGET_ONTOLOGY %>% filter(Drugs != "", !is.na(Drugs)) %>%  dplyr::select(TTDID, Drugs, UniProtID, Pathway) %>% distinct()
TARGET_ONTOLOGY <- as_tibble(apply(TARGET_ONTOLOGY,2,function(x) trimws(toupper(x))))
TARGET_ONTOLOGY            <- parseColByIdx(TARGET_ONTOLOGY,"Drugs")
TARGET_ONTOLOGY            <- parseColByIdx(TARGET_ONTOLOGY,"UniProtID")
TARGET_ONTOLOGY            <- parseColByIdx(TARGET_ONTOLOGY,"Pathway"1)

# transform pathway to 
# pathways, load and preprocess
TARGET2KEGGPATHWAYS <- tbl(src,"REFERENCE.TARGET2KEGGPATHWAYS") %>% filter(TTDID != "") %>%  collect()
TARGET2KEGGPATHWAYS <- TARGET2KEGGPATHWAYS %>% dplyr::rename(PATH_ID = KEGG_PATHWAYID, PATH_NAME = KEGG_PATHWAY_NAME)
human               <- grepl("hsa",TARGET2KEGGPATHWAYS$PATH_ID)
TARGET2KEGGPATHWAYS <- TARGET2KEGGPATHWAYS[human,]

TARGET2WIKIPATHWAYS <- tbl(src,"REFERENCE.TARGET2WIKIPATHWAYS") %>%   collect()
TARGET2WIKIPATHWAYS <- TARGET2WIKIPATHWAYS %>% dplyr::rename(PATH_ID = WIKI_PATHWAY_ID, PATH_NAME =  WIKI_PATHWAY_NAME)
PATHWAYS            <- rbind(TARGET2KEGGPATHWAYS,TARGET2WIKIPATHWAYS) %>% distinct()
PATHWAYS$PATH_NAME  <- trimws(toupper(PATHWAYS$PATH_NAME))

# match TARGET_ONTOLOGY$Pathway to PATHWAYS$PATH_NAME and retrieve PATHWAYS$PATH_ID
TARGET_ONTOLOGY.P1 <- 
  inner_join(
    TARGET_ONTOLOGY, 
    PATHWAYS %>% dplyr::select(-PATH_NAME), 
    by = "TTDID") %>% 
  dplyr::select(-Pathway) %>% 
  distinct()

TARGET_ONTOLOGY.P2 <- 
  inner_join(
    TARGET_ONTOLOGY, 
    PATHWAYS %>% dplyr::select(-TTDID), 
    by = c("Pathway"="PATH_NAME")) %>%
  dplyr::select(-Pathway) %>% 
  distinct()

TARGET_ONTOLOGY <- rbind(TARGET_ONTOLOGY.P1, TARGET_ONTOLOGY.P2) %>% distinct() %>% dplyr::rename(PATH = PATH_ID)
```

## DRD.W3.1 (enlarged namespace) + TARGET_ONTOLOGY -> DRD.W3.1.x 
```{r}

DRD.W3.1 <- left_join(DRD.W3.1e, TARGET_ONTOLOGY, by = c("NAME"="Drugs")) %>% dplyr::select(-NAME) %>% distinct()
DRD.W3.1 <- rbind(
  DRD.W3.1 %>% dplyr::select(-UniProt.ID) %>% distinct(),
  DRD.W3.1 %>% dplyr::select(-UniProtID) %>% dplyr::rename(UniProtID = UniProt.ID) %>% distinct()) %>% 
  distinct()

# add TTDID Pathways
####################
DRD.W3.2 <- left_join(DRD.W3.1, PATHWAYS %>% dplyr::select(-PATH_NAME) %>% distinct()) %>% 
  dplyr::select(-TTDID) %>% distinct()

DRD.W3.3 <- rbind(
  DRD.W3.2 %>% dplyr::select(-PATH_ID) %>% distinct(),
  DRD.W3.2 %>% dplyr::select(-PATH) %>% distinct() %>% dplyr::rename(PATH = PATH_ID)
) %>% distinct()

# postprocessing
################
x        <- which(is.na(DRD.W3.3$UniProtID) & is.na(DRD.W3.3$PATH))
DRD.W3.3 <- DRD.W3.3[-x,]
```


# module UNIPROT

```{r}
columns(org.Hs.eg.db)
u <- unique(DRD.W3.3$UniProtID)

# UNIPROT to GO
UNIPROT2GO       <- as_tibble(AnnotationDbi::select(org.Hs.eg.db, keys = u, keytype="UNIPROT" ,columns=c("UNIPROT","GO")))
UNIPROT2GO       <- UNIPROT2GO %>% dplyr::select(-EVIDENCE) %>% distinct()
UNIPROT2GO$GO    <- gsub("GO:","",UNIPROT2GO$GO)
UNIPROT2GO       <- UNIPROT2GO[!is.na(UNIPROT2GO$GO),]
UNIPROT2GO$GO    <- paste(UNIPROT2GO$ONTOLOGY,".",UNIPROT2GO$GO,sep="")
UNIPROT2GO       <- UNIPROT2GO %>% dplyr::select(-ONTOLOGY) %>% distinct()

# UNIPROT to KEGG PATHWAYS
UNIPROT2KEGG <- as_tibble(AnnotationDbi::select(org.Hs.eg.db, keys = u, keytype="UNIPROT" ,columns=c("UNIPROT","PATH")))
UNIPROT2KEGG <- UNIPROT2KEGG[complete.cases(UNIPROT2KEGG),] %>% distinct()
UNIPROT2KEGG$PATH <- sapply(X = UNIPROT2KEGG$PATH, FUN = function(x) paste("hsa",x,sep=""))

# UNIPROT to ENTREZ
UNIPROT2ENTREZ <- as_tibble(AnnotationDbi::select(org.Hs.eg.db, keys = u, keytype="UNIPROT" ,columns=c("UNIPROT","SYMBOL")))
UNIPROT2ENTREZ <- UNIPROT2ENTREZ[complete.cases(UNIPROT2ENTREZ),] %>% distinct() %>% dplyr::rename(GENESYMBOL = SYMBOL)
```

 
```{r}
# join GO, KEGGPATH, ENTREZ by UNIPROT to DRD.W3.3
##################################################

# go
DRD.W3.4   <- left_join(DRD.W3.3, UNIPROT2GO, by = c("UniProtID" = "UNIPROT")) %>% distinct()

# kegg
DRD.W3.4.x <- DRD.W3.4 %>% dplyr::select(-PATH) %>% distinct()
DRD.W3.4.x <- left_join(DRD.W3.4.x, UNIPROT2KEGG, by = c("UniProtID" = "UNIPROT"))
DRD.W3.4   <- rbind(DRD.W3.4, DRD.W3.4.x) %>% distinct()

# entrez
DRD.W3.4 <- left_join(DRD.W3.4, UNIPROT2ENTREZ, by = c("UniProtID" = "UNIPROT"))


```

# WAY 4

## TARGET2UNIPROT, parse UNIPROT_ID
```{r}
TARGET2UNIPROT <- tbl(src,"REFERENCE.TARGET2UNIPROT") %>% dplyr::select(TTD_TARGET_ID, UNIPROT_ID) %>% 
  filter(UNIPROT_ID != "", !is.na(UNIPROT_ID)) %>% distinct() %>% collect()
TARGET2UNIPROT$UNIPROT_ID <- gsub("[\\(].+[\\)]","",TARGET2UNIPROT$UNIPROT_ID)
TARGET2UNIPROT            <- parseColByIdx(TARGET2UNIPROT,"UNIPROT_ID")

```

## TARGET2DISEASES, parse ICD9
```{r}
TARGET2DISEASES <- tbl(src,"REFERENCE.TARGET2DISEASES") %>% dplyr::select(TTDTARGETID, ICD9) %>% distinct() %>% collect()
TARGET2DISEASES$ICD9[TARGET2DISEASES$ICD9 == "."] <- NA
del <- c("585.9585.1-585.5403","280-285, 585.9585.1-585.5403")
TARGET2DISEASES <- TARGET2DISEASES %>% filter(ICD9 != "", !is.na(ICD9), !(ICD9 %in% del))
TARGET2DISEASES <- ParseICD9(TARGET2DISEASES %>% dplyr::rename(DIAG = ICD9)) %>% dplyr::rename(ICD9 = DIAG)
# u <- unique(as.character(TARGET2DISEASES$ICD9))
# parsed <- as.data.frame(parseICD9(u))
# head(parsed)
# colnames(parsed) <- c("ICD9","ICD9.parsed")
# TARGET2DISEASES <- left_join(TARGET2DISEASES, parsed) %>% dplyr::select(-ICD9) %>% dplyr::rename(ICD9 = ICD9.parsed) %>% distinct()
# TARGET2DISEASES$ICD9 <- trimws(TARGET2DISEASES$ICD9)
```

## DRD <-> TARGET2UNIPROT <-> TARGET2DISEASES -> DRD
```{r}

DRD.W4.1 <- left_join(DRD.W3.4,TARGET2UNIPROT,by=c("UniProtID"="UNIPROT_ID")) %>% dplyr::rename(TTDTARGETID = TTD_TARGET_ID)
DRD.W4.2 <- left_join(DRD.W4.1,TARGET2DISEASES)
DRD.W4.2 <- DRD.W4.2 %>% dplyr::select(-UniProtID, -TTDTARGETID) %>% distinct()

```

## add PHEWAS_STRING to DRD.W4.2
```{r}
DRD.W4.3 <- left_join(DRD.W4.2, ICD9GROUPING %>% dplyr::select(ICD9, PHEWAS_CODE) %>% distinct())
DRD.W4.3 <-
  DRD.W4.3 %>% dplyr::rename(
  INDC.ICD9 = ICD9,
  INDC.PHEWAS = PHEWAS_CODE,
  TARGET.PATH = PATH,
  TARGET.GO = GO,
  TARGET.GENE = GENESYMBOL
  )
```

# Interface I1

## DRD.W1.3 + DRD.W2.6 -> DRD.W2.7 
```{r}

# DRD.W1.3 + DRD.W2.6 -> DRD.W2.7
#################################

# spread DRD.W1.3
DRD.W1.3.GO   <- as_tibble(cbind(DRD.W1.3 %>% dplyr::select(TTDDRUGID, INDC.GO) %>% distinct(), val = TRUE)) 
DRD.W1.3.PATH <- as_tibble(cbind(DRD.W1.3 %>% dplyr::select(TTDDRUGID, INDC.PATH) %>% distinct(), val = TRUE)) 

DRD.W1.3.GO   <- spread(DRD.W1.3.GO, key = INDC.GO, value = val, fill = FALSE, sep = "_x_")
DRD.W1.3.PATH <- spread(DRD.W1.3.PATH, key = INDC.PATH, value = val, fill = FALSE, sep = "_x_")

DRD.W1.3.spread <- cbind(DRD.W1.3.PATH, DRD.W1.3.GO[,-1])
rm(DRD.W1.3.GO, DRD.W1.3.PATH )

# spread DRD.W2.6
DRD.W2.6.spread <- spread(DRD.W2.6, key = MEDDRA, value = WHO, fill = FALSE, sep = "_x_")

# -> DRD.W2.7
DRD.W2.7 <- as_tibble(unique(c(DRD.W1.3.spread$TTDDRUGID, DRD.W2.6.spread$TTDDRUGID)))
colnames(DRD.W2.7) <-  "TTDDRUGID"
DRD.W2.7 <- left_join(DRD.W2.7, DRD.W1.3.spread, by = "TTDDRUGID")
DRD.W2.7 <- left_join(DRD.W2.7, DRD.W2.6.spread, by = "TTDDRUGID")

```

## transform DRD.W.4.3 

In addition add "DRUG Indication PHEWAS Codes"" from DRD.W0.2 to
the "Target Indication PHEWAS Codes" from DRD.W4.3
```{r}

# spread DRD.W4.3 piecewise
###########################

# path
DRD.W4.3.PATH   <- DRD.W4.3 %>% dplyr::select(TTDDRUGID, TARGET.PATH) %>% distinct()
DRD.W4.3.PATH   <- DRD.W4.3.PATH[!is.na(DRD.W4.3.PATH$TARGET.PATH),]
DRD.W4.3.PATH   <- cbind(DRD.W4.3.PATH, val = TRUE)
DRD.W4.3.PATH   <- spread(DRD.W4.3.PATH, TARGET.PATH, val, sep = "_x_", fill = FALSE)

# go
DRD.W4.3.GO   <- DRD.W4.3 %>% dplyr::select(TTDDRUGID, TARGET.GO) %>% distinct()
DRD.W4.3.GO   <- DRD.W4.3.GO[!is.na(DRD.W4.3.GO$TARGET.GO),]
DRD.W4.3.GO   <- cbind(DRD.W4.3.GO, val = TRUE)
DRD.W4.3.GO   <- spread(DRD.W4.3.GO, TARGET.GO, val, sep = "_x_", fill = FALSE)

# genes
DRD.W4.3.GENE <- DRD.W4.3 %>% dplyr::select(TTDDRUGID, TARGET.GENE ) %>% distinct()
DRD.W4.3.GENE$TARGET.GENE  <- as.character(DRD.W4.3.GENE$TARGET.GENE  )
DRD.W4.3.GENE   <- cbind(DRD.W4.3.GENE, val = TRUE)
DRD.W4.3.GENE   <- spread(DRD.W4.3.GENE, TARGET.GENE, val, sep = "_x_", fill = FALSE)
# phewas
########

# phewas
DRD.W4.3.PHEWAS <- DRD.W4.3 %>% dplyr::select(TTDDRUGID, INDC.PHEWAS) %>% distinct()
DRD.W4.3.PHEWAS$INDC.PHEWAS <- as.character(DRD.W4.3.PHEWAS$INDC.PHEWAS )

# # phewas from DRD.W0.2
DRD.W0.2.PHEWAS <-
  DRD.W0.2 %>%  dplyr::select(TTDDRUGID, PHEWAS_CODE) %>% 
  dplyr::rename(INDC.PHEWAS = PHEWAS_CODE) %>% 
  filter(!is.na(INDC.PHEWAS)) %>%  distinct() 

DRD.W0.2.PHEWAS$INDC.PHEWAS <- as.character(DRD.W0.2.PHEWAS$INDC.PHEWAS)
DRD.W4.3.PHEWAS             <- rbind(DRD.W4.3.PHEWAS, DRD.W0.2.PHEWAS) %>% distinct()

# #
DRD.W4.3.PHEWAS <- DRD.W4.3.PHEWAS[!is.na(DRD.W4.3.PHEWAS$INDC.PHEWAS),]
DRD.W4.3.PHEWAS <- cbind(DRD.W4.3.PHEWAS, val = TRUE)
DRD.W4.3.PHEWAS$INDC.PHEWAS <- floor(10*as.numeric(DRD.W4.3.PHEWAS$INDC.PHEWAS)) / 10 
DRD.W4.3.PHEWAS <- DRD.W4.3.PHEWAS %>% distinct()
# # round to highlevel phewas

DRD.W4.3.PHEWAS <- spread(DRD.W4.3.PHEWAS, INDC.PHEWAS, val, sep = "_x_", fill = FALSE)

# # only for heatmap, comment after usage !
# DRD.W4.3.PATH   <- DRD.W4.3 %>% dplyr::select(TTDDRUGID, TARGET.PATH) %>% distinct()
# DRD.W4.3.PATH   <- DRD.W4.3.PATH[!is.na(DRD.W4.3.PATH$TARGET.PATH),]
# DRD.W4.3.PATH   <- cbind(DRD.W4.3.PATH, val = 1)
# DRD.W4.3.PATH   <- spread(DRD.W4.3.PATH, TARGET.PATH, val, sep = "_x_", fill = 0)
# heatmap.2(apply(DRD.W4.3.PATH[,-1],2,as.numeric), dendrogram = 'row',trace = 'none', xlab = "PATH")
# 
# DRD.W4.3.PHEWAS <- DRD.W4.3 %>% dplyr::select(TTDDRUGID, INDC.PHEWAS) %>% distinct()
# DRD.W4.3.PHEWAS <- DRD.W4.3.PHEWAS[!is.na(DRD.W4.3.PHEWAS$INDC.PHEWAS),]
# DRD.W4.3.PHEWAS <- cbind(DRD.W4.3.PHEWAS, val = 1)
# DRD.W4.3.PHEWAS <- spread(DRD.W4.3.PHEWAS, INDC.PHEWAS, val, sep = "_x_", fill = 0)
# heatmap.2(apply(DRD.W4.3.PHEWAS[,-1],2,as.numeric), dendrogram = 'row',trace = 'none', xlab = "PATH")

# join spreaded tables
DRD.W4.4 <- as_tibble(unique(c(DRD.W4.3.PATH$TTDDRUGID, DRD.W4.3.PHEWAS$TTDDRUGID)))
colnames(DRD.W4.4) <- "TTDDRUGID"
DRD.W4.4 <- left_join(DRD.W4.4, DRD.W4.3.PATH, by = "TTDDRUGID")
DRD.W4.4 <- left_join(DRD.W4.4, DRD.W4.3.PHEWAS, by = "TTDDRUGID")
DRD.W4.4 <- left_join(DRD.W4.4, DRD.W4.3.GO, by = "TTDDRUGID")
DRD.W4.4 <- left_join(DRD.W4.4, DRD.W4.3.GENE, by = "TTDDRUGID")
rm(DRD.W4.3.PATH, DRD.W4.3.PHEWAS, DRD.W0.2.PHEWAS, DRD.W4.3.GO, DRD.W4.3.GENE)
```

# WAY 5

## combine DRD.W4.4 + DRD.W2.7 to DRD.W5.2
```{r}

DRD.W5.2 <- as_tibble(unique(c(DRD.W4.4$TTDDRUGID, DRD.W2.7$TTDDRUGID)))
colnames(DRD.W5.2) <- "TTDDRUGID"
DRD.W5.2 <- left_join(DRD.W5.2, DRD.W4.4, by = "TTDDRUGID")
DRD.W5.2 <- left_join(DRD.W5.2, DRD.W2.7, by = "TTDDRUGID")
```

# WAY 6

## combine DRD.W5.2 + DRD.W0.2

Join PATIENTS-table with the remaining drug-descriptor features, with exception
to DRD.W0.2$PHEWAS_CODE, because that's already been done in DRD.W4.4

```{r}

# preprocessing DRD.W0.2
########################
DRD.W0.2.x <- DRD.W0.2 %>% dplyr::select(-ICD9, -PHEWAS_CODE, -DRUGNAME) %>% distinct()

# drop redundant rows with empty PHARMCLS1
x1 <- 
  DRD.W0.2.x %>% 
  filter(PHARMCLS1 == "") %>% 
  dplyr::select(SUBSTANCENAME) %>% 
  distinct()

x2 <- 
  DRD.W0.2.x %>% 
  filter(SUBSTANCENAME %in% x1$SUBSTANCENAME) %>% 
  group_by(SUBSTANCENAME) %>% 
  dplyr::select(PHARMCLS1) %>% 
  tally() %>% 
  filter(n>1)

x3 <- 
  DRD.W0.2.x %>% 
  filter(SUBSTANCENAME %in% x2$SUBSTANCENAME) %>% 
  dplyr::select(SUBSTANCENAME, TTDDRUGID) %>% 
  distinct() %>% 
  group_by(SUBSTANCENAME) %>% 
  tally() %>% 
  filter(n==1)

DRD.W0.2.x <-
  DRD.W0.2.x %>% 
  filter(!(SUBSTANCENAME %in% x3$SUBSTANCENAME & PHARMCLS1 == ""))

# replace emtpy PHARMCLS1 with NA
DRD.W0.2.x$PHARMCLS1[DRD.W0.2.x$PHARMCLS1 == ""] <- NA


# join other drug descriptors
#############################
DRD.W6.1 <- inner_join(DRD.W0.2.x, DRD.W5.2,by = "TTDDRUGID")
DRD.W6.1[is.na(DRD.W6.1)] <- FALSE

# if one SUBSTANCENAME has many TTDDRUGID, then combine features
#################################################################
DRD.W6.2 <- DRD.W6.1 %>% dplyr::select(-TTDDRUGID) %>% distinct()
DRD.W6.3 <- aggregate( x = DRD.W6.2, by = list(DRD.W6.2$SUBSTANCENAME, DRD.W6.2$PHARMCLS1),FUN=max, na.rm=TRUE)

DRD.W6.3 <- DRD.W6.3 %>% dplyr::select(-Group.1, -Group.2)
```

# save feature matrix
```{r}
DRUG.FEATMAT <- DRD.W6.3
save(DRUG.FEATMAT,file = "atogQueries/DRUG.FEATMAT.RData",compress = T)

```

