---
title: "descriptive statistics of adult (AGE 18-66)"
subtitle: "re-do epd183"
author: "Thomas Gerlach"
output: html_notebook
date: "2017 APR 05"
---

# setup global variables
```{r, echo = FALSE, include = FALSE}
rm(list=setdiff(ls(),"all_events.icd9"))
library(data.table)
knitr::opts_chunk$set(warning = FALSE, fig.width = 10,fig.height = 7)

source("~/epilepsie_preprocessing_descrStat/z_start.R") # initalize netezza DB, load core data

params <- vector("list")
params$incidence.buffer <- 180
params$EPILEPSY <-  c("Convulsions","Epilepsy, recurrent seizures, convulsions","Partial epilepsy", "Epilepsy", "Generalized convulsive epilepsy" )

if(!exists("all_events.icd9")) load("~/epilepsie_preprocessing_descrStat/data/re-do_epd183/all_events.icd9.RData")

PATIENTS_AGE_REGION <-  as.data.frame(all_events.icd9$data)
```  

# filter PATIENTS 
```{r, include = FALSE}

nPatObsCounter <- function(y) {
  list(addCount=
         function(x,descr) {
           nAdult <- x %>% filter(AGE >= 18) %>% dplyr::select(ENROLID) %>% distinct() %>% tally()
           nChild <- x %>% filter(AGE <= 16) %>% dplyr::select(ENROLID) %>% distinct() %>% tally()
           nObs   <- nrow(x)
           xx <- cbind(descr, nObs, nAdult, nChild)
           colnames(xx) <- c("description","nObs", "Adult, >= 18", "Child, <= 16")
           y <<- rbind(y,xx)
           },
       value = function() return(y)
  )
}

source("~/epilepsie_preprocessing_descrStat/functions/n_patient_per_period.R")
source("~/epilepsie_preprocessing_descrStat/functions/add_Phew_PhewHighlevel.R")


# filter
########
patObsCount <- nPatObsCounter(NULL)
patObsCount$addCount(PATIENTS_AGE_REGION,"unconstraint")

# Keep only patients within valid time period 
PATIENTS_AGE_REGION <- PATIENTS_AGE_REGION %>% filter(has2yMH == TRUE)
patObsCount$addCount(PATIENTS_AGE_REGION,"2y MH + 1 quarter FU")

# adult only (AGE 18-66)
PATIENTS_AGE_REGION <- PATIENTS_AGE_REGION %>% filter(AGE >= 18) %>% distinct()
patObsCount$addCount(PATIENTS_AGE_REGION,"AGE 18-66 ")

# only patients which were treated with AEDs in their medical history
pat.with.aeds <- PATIENTS_AGE_REGION %>% filter(isAED == TRUE, IXDAYS <= 180 & SUBSTANCENAME != "") %>% select(ENROLID) %>% distinct()
PATIENTS_AGE_REGION <-   PATIENTS_AGE_REGION %>% filter(ENROLID %in% pat.with.aeds$ENROLID)
patObsCount$addCount(PATIENTS_AGE_REGION,"AED treatment")

## add main comorbidities
PATIENTS_AGE_REGION <- AddMainComorb(PATIENTS_AGE_REGION, cohort = "adult")
PATIENTS_AGE_REGION <- unique(PATIENTS_AGE_REGION)
```

# aggregate AED substancenames
```{r}
PATIENTS_AGE_REGION %>% filter(isAED == TRUE) %>% select(SUBSTANCENAME) %>% distinct()
```


# print MAIN.COMORB to PHEWAS_STRING assignment
```{r}
MAIN.COMORB <- (MainComorbAdult())[[1]]
MAIN.COMORB.PHEWAS <- (MainComorbAdult())[[2]]
knitr::kable(MAIN.COMORB.PHEWAS)
```


# ECDF of first occurrence of a main-comorbidity
```{r, fig.height=7, fig.width=7}
main.comorb.first.ixday <-
  PATIENTS_AGE_REGION %>% 
  filter(MAIN.COMORB != "") %>% 
  dplyr::select(ENROLID, IXDAYS, MAIN.COMORB) %>% 
  distinct()

main.comorb.first.ixday <- aggregate(IXDAYS ~., data = main.comorb.first.ixday , FUN = min)

ggplot(main.comorb.first.ixday, aes(IXDAYS)) +  
  facet_wrap( ~ MAIN.COMORB) +
  stat_ecdf() + 
  ggtitle(paste("ECDF of first occurrence of diagnosis (PHEWAS) \n aggregated by main-comorbidity \n blue line = 180 days, red lines = 80%, 90%")) +
  geom_vline(xintercept = 180, colour = "blue", linetype = "longdash", size = 0.1) +
  geom_hline(yintercept = c(0.8, 0.9), colour = "red", linetype = "longdash", size=0.1) +
  theme(panel.background = element_rect(fill = "grey"))

```

# ECDF of first treatment with an AED
```{r}

PATIENTS_AGE_REGION <- KeepIncidentAndCensored(PATIENTS_AGE_REGION)
n_distinct(PATIENTS_AGE_REGION$ENROLID)

firstAEDTreatment <- 
  PATIENTS_AGE_REGION %>% 
  filter(isAED == TRUE) %>% 
  select(ENROLID, IXDAYS) %>% 
  distinct() %>% 
  group_by(ENROLID) %>% mutate(firstAEDTreatment = min(IXDAYS)) %>% 
  ungroup() %>% dplyr::select(-IXDAYS) %>% distinct()

head(firstAEDTreatment)
n_distinct(firstAEDTreatment$ENROLID)

ggplot(firstAEDTreatment, aes(firstAEDTreatment)) +  
  stat_ecdf() + 
  ggtitle(paste("ECDF of first treatment with an AED \n blue line = -730, 180 days, red lines = 80%, 90%")) +
  geom_vline(xintercept = c(-730,180), colour = "blue", linetype = "longdash", size = 0.1) +
  geom_hline(yintercept = c(0.8, 0.9), colour = "red", linetype = "longdash", size=0.1)

summary(firstAEDTreatment$firstAEDTreatment)
```

# MAIN.COMORB
## prevalence
### by PHEWAS_STRING
```{r, warning=FALSE}

# without any main-comorbidity
x <- 
  PATIENTS_AGE_REGION %>% 
  filter(!is.na(MAIN.COMORB)) %>% 
  dplyr::select(ENROLID) %>%
  distinct()

  # with an follow up of at least 2 quarters ( 183 days = ceiling(365/4 * 2) )
  x <- 
    PATIENTS_AGE_REGION %>% 
    filter(!(ENROLID %in% x$ENROLID)) %>% 
    select(ENROLID, IXDAYS) %>% 
    distinct() %>% 
    group_by(ENROLID) %>% 
    mutate(maxIXDAYS = max(IXDAYS)) %>% filter(maxIXDAYS >= 183) %>% 
    select(-IXDAYS) %>% 
    distinct()

patObsCount$addCount(PATIENTS_AGE_REGION %>% filter(ENROLID %in% x$ENROLID),paste("without any main-comorbidity & >= 2q FU"))

# with main co-morbidities
x <- 
  main.comorb.first.ixday %>% 
  filter(IXDAYS >= params$incidence.buffer) %>% 
  dplyr::select(ENROLID) %>%
  distinct()

patObsCount$addCount(
  PATIENTS_AGE_REGION %>% filter(ENROLID %in% x$ENROLID),
  paste(
  "incident main-comorbidities >= ",
  params$incidence.buffer,
  " days"
  )
  )

## by each main-comorbidity
main.comorb.first.ixday.inc <- main.comorb.first.ixday %>% filter(IXDAYS >= 365/4*2) %>% distinct()
for (m in MAIN.COMORB) {
  m.enrolids <-
    main.comorb.first.ixday.inc %>%
    filter(MAIN.COMORB == m) %>% 
    dplyr::select(ENROLID) %>% 
    distinct()
  m.enrolids <- m.enrolids$ENROLID
  
  patObsCount$addCount(
  PATIENTS_AGE_REGION %>% filter(ENROLID %in% m.enrolids),
  paste("-",m)
  )
}


# n patients per MAIN.COMORB, by PHEWAS_STRING
##############################################
MAIN.COMORB.PHEWAS <-
  left_join(
  MAIN.COMORB.PHEWAS,
  PATIENTS_AGE_REGION %>%
  filter(PHEWAS_STRING %in% MAIN.COMORB.PHEWAS$PHEWAS_STRING) %>%
  dplyr::select(ENROLID, PHEWAS_STRING) %>%
  distinct() %>%
  group_by(PHEWAS_STRING) %>%
  tally(),
  by = "PHEWAS_STRING"
  )

knitr::kable(MAIN.COMORB.PHEWAS)
```

### by MAIN.COMORB
```{r, echo = FALSE}
# count prevalent patients per MC
prev.allAEDs <- 
  PATIENTS_AGE_REGION %>% 
  filter(MAIN.COMORB != "") %>% 
  dplyr::select(ENROLID, MAIN.COMORB) %>% 
  distinct() %>% 
  group_by(MAIN.COMORB) %>% 
  tally() %>% 
  dplyr::rename(n.prevalent = n)

knitr::kable(prev.allAEDs[order(-prev.allAEDs$n.prevalent),])

cat("The sum of patients within a comorbidity is NOT the sum of patients in the PHEWAS_STRING by MAIN.COMORB due to overlaps.")
```

## incidence
```{r, warning=FALSE}
# 2. plot number of patients with incident main-comorbidity occurrence as a function of <buffer.int>-time
#########################################################################################################
buffer.int <- 1 : max(main.comorb.first.ixday$IXDAYS)

# summarised over all patients
##############################
x <- 
  unlist(
    mapply(
      function(x) main.comorb.first.ixday %>% filter(IXDAYS >= x) %>% dplyr::select(ENROLID) %>% distinct() %>% tally(),
      buffer.int)
    )

all.incident <- as.data.frame(cbind(buffer.int, x))

ggplot(all.incident, aes(buffer.int, x)) +
  geom_line(aes(y = x, colour = "sum") ) +
  ggtitle("Number of patients with new main-comorbidities at time >= IXDAYS \n summarized over all main-comorbidities") +
  xlab("IXDAYS") +
  theme(panel.background = element_rect(fill = "grey"))

# by main-comorbidity
#####################
first.ixdays.vs.buffer <- as.data.frame(matrix(data = NA, nrow = length(buffer.int), ncol = length(MAIN.COMORB)))
colnames(first.ixdays.vs.buffer) <- MAIN.COMORB
first.ixdays.vs.buffer <- cbind(buffer = buffer.int, first.ixdays.vs.buffer)

for( m in unique(main.comorb.first.ixday$MAIN.COMORB)) {
  first.ixdays.vs.buffer[, colnames(first.ixdays.vs.buffer) == m] <-
    unlist(
      mapply(
        function(x)
          main.comorb.first.ixday %>% filter(MAIN.COMORB == m, IXDAYS >= x) %>% dplyr::select(ENROLID) %>% distinct() %>% tally(),
        x = buffer.int))
}


ggplot(first.ixdays.vs.buffer, aes(buffer.int)) +
  geom_line(aes(y = Hypertension, colour = "Hypertension") ) +
  geom_line(aes(y= Diabetes, colour = "Diabetes")) +
  geom_line(aes(y= Hyperlipidemia, colour =  "Hyperlipidemia")) +
  geom_line(aes(y= Anxiety, colour = "Anxiety")) + 
  geom_line(aes(y= Stroke.IschemAttack, colour = "Stroke, ischemic attack")) +
  geom_line(aes(y= Migraine, colour = "Migraine")) +
  geom_line(aes(y= Depression, colour = "Depression")) +
  geom_line(aes(y= Overweight, colour = "Overweight")) +
  geom_line(aes(y= Bipolar.Schizophrenia, colour = "Bipolar.Schizophrenia")) +
  ggtitle("Number of patients with new main-comorbidities at time >= IXDAYS \n by main-comorbidity") +
  xlab("IXDAYS") +
  theme(panel.background = element_rect(fill = "grey"))

mc.inc180 <- 
  first.ixdays.vs.buffer %>%
  filter(buffer == 180) %>% 
  dplyr::select(-buffer) %>%
  t() 

mc.inc180 <- data.frame(MC = row.names(mc.inc180), n.incident = mc.inc180)
mc.inc180 <- mc.inc180[order(-mc.inc180$n.incident),]
row.names(mc.inc180) <- c()
knitr::kable(mc.inc180 )
```

### incidence in % of population by time
```{r}
n_distinct(PATIENTS_AGE_REGION$ENROLID) # (patients with AED treatment)

# get max IXDAY per patient
pat.maxIxdays <- 
  PATIENTS_AGE_REGION %>% dplyr::select(ENROLID, IXDAYS) %>% distinct() %>% 
  group_by(ENROLID) %>% mutate(maxIXDAYS = max(IXDAYS)) %>% ungroup() %>% 
  dplyr::select(-IXDAYS) %>% distinct()

# plot number of patients by time
min(pat.maxIxdays$maxIXDAYS)
t <- 0:max(pat.maxIxdays$maxIXDAYS)
t.pop <- sapply(t, function(x) sum(pat.maxIxdays$maxIXDAYS >= x))
t.pop <- data.frame(IXDAYS = t, popsize = t.pop)

ggplot(t.pop, aes(IXDAYS, popsize)) +
  geom_line() +
  ggtitle("popsize by time") +
  xlab("IXDAYS") +
  geom_vline(xintercept = 90, linetype = "longdash") +
  theme(panel.background = element_rect(fill = "grey"))

# plot percent of incident patients relative to total patients available at that time
all.incident.normalized <- all.incident %>% dplyr::rename(n.incident = x)
row.names(all.incident.normalized) <- c()
all.incident.normalized <- left_join(all.incident.normalized, t.pop, by = c("buffer.int" = "IXDAYS"))
all.incident.normalized <- all.incident.normalized %>% mutate(n.incident.per1000 = n.incident / popsize)

ggplot(all.incident.normalized, aes(buffer.int, n.incident.per1000)) +
  geom_line(aes(colour = "fraction"))  +
  ggtitle("incidence in % of population by time \n all main-comorbidities") +
  xlab("IXDAYS") +
  geom_vline(xintercept = 90, linetype = "longdash") +
  theme(panel.background = element_rect(fill = "grey"))

# plot percent of incident patients relative to total patients available at that time
# by MC
first.ixdays.vs.buffer.normalized <- first.ixdays.vs.buffer %>% left_join(t.pop,  by = c("buffer" = "IXDAYS"))
first.ixdays.vs.buffer.normalized[,-c(1,ncol(first.ixdays.vs.buffer.normalized))] <- 
  first.ixdays.vs.buffer.normalized[,-c(1,ncol(first.ixdays.vs.buffer.normalized))] / first.ixdays.vs.buffer.normalized$popsize

ggplot(first.ixdays.vs.buffer.normalized, aes(buffer.int)) +
  geom_line(aes(y = Hypertension, colour = "Hypertension") ) +
  geom_line(aes(y= Diabetes, colour = "Diabetes")) +
  geom_line(aes(y= Hyperlipidemia, colour =  "Hyperlipidemia")) +
  geom_line(aes(y= Anxiety, colour = "Anxiety")) + 
  geom_line(aes(y= Stroke.IschemAttack, colour = "Stroke, ischemic attack")) +
  geom_line(aes(y= Migraine, colour = "Migraine")) +
  geom_line(aes(y= Depression, colour = "Depression")) +
  geom_line(aes(y= Overweight, colour = "Overweight")) +
  geom_line(aes(y= Bipolar.Schizophrenia, colour = "Bipolar.Schizophrenia")) +
  ggtitle("incidence in % of population by time \n by main-comorbidity") +
  xlab("IXDAYS") +
  theme(panel.background = element_rect(fill = "grey"))

```


# kaplan meyer plot for main-comorbidity events
```{r, fig.height=10, fig.width=10}
# get max observation date for each patient
# pat.maxIxdays
# for patients with an MC, get event date for incident main-comorbidity
# main.comorb.first.ixday
firstIxday.byMC <- dcast(main.comorb.first.ixday %>% filter(IXDAYS >= 365/4*2), ENROLID ~ MAIN.COMORB, value.var = "IXDAYS")

list.survival <- vector("list",length(MAIN.COMORB))
names(list.survival) <- MAIN.COMORB

# 
for (m in MAIN.COMORB) {
  m.event <- firstIxday.byMC[,colnames(firstIxday.byMC) %in% c("ENROLID",m)] # retrieve events for MC "m"
  m.event <- m.event[complete.cases(m.event),] 
  colnames(m.event) <- c("ENROLID","time")
  m.event$event <- 1 # 1 flags "the event"
  m.censored <- pat.maxIxdays %>% filter(!(ENROLID %in% m.event$ENROLID)) # drop patients with event
  colnames(m.censored) <- c("ENROLID","time")
  m.censored$event <- 0 # 0 flags censored
  # bind event + censored + split to "survival table"
  m.surv <- data.frame(rbind(m.event, m.censored))
  m.surv$MC <- m
  l.idx <- which(names(list.survival) %in% m)
  list.survival[[l.idx]] <- m.surv
}


df.survival <- do.call(rbind, lapply(list.survival, data.frame, stringsAsFactors=FALSE)) # list to df
library(survival)
(KM <- survfit(Surv(time, event) ~ MC, type="kaplan-meier", conf.type="log", data=df.survival))

survminer::ggsurvplot(KM,pval = TRUE, risk.table = TRUE, ylim = c(0.5,1), main = "kaplan meier estimates of incident main-comorbidities \n adult")
```



# kaplan meyer plot for main-comorbidity events, for model-population
```{r, fig.height=10, fig.width=10}
# get max observation date for each patient
# pat.maxIxdays
# for patients with an MC, get event date for incident main-comorbidity
# main.comorb.first.ixday
load("~/epilepsie_preprocessing_descrStat/output/featureMatrix/censored.incident.RData")
firstIxday.byMC <- dcast(main.comorb.first.ixday %>% filter(ENROLID %in% censored.incident$ENROLID[censored.incident$status != "censored"]), ENROLID ~ MAIN.COMORB, value.var = "IXDAYS")

list.survival <- vector("list",length(MAIN.COMORB))
names(list.survival) <- MAIN.COMORB

# 
for (m in MAIN.COMORB) {
  m.event <- firstIxday.byMC[,colnames(firstIxday.byMC) %in% c("ENROLID",m)] # retrieve events for MC "m"
  m.event <- m.event[complete.cases(m.event),] 
  colnames(m.event) <- c("ENROLID","time")
  m.event$event <- 1 # 1 flags "the event"
  m.censored <- pat.maxIxdays %>% filter(!(ENROLID %in% m.event$ENROLID)) # drop patients with event
  colnames(m.censored) <- c("ENROLID","time")
  m.censored$event <- 0 # 0 flags censored
  # bind event + censored + split to "survival table"
  m.surv <- data.frame(rbind(m.event, m.censored))
  m.surv$MC <- m
  l.idx <- which(names(list.survival) %in% m)
  list.survival[[l.idx]] <- m.surv
}


df.survival <- do.call(rbind, lapply(list.survival, data.frame, stringsAsFactors=FALSE)) # list to df
library(survival)
(KM <- survfit(Surv(time, event) ~ MC, type="kaplan-meier", conf.type="log", data=df.survival))

survminer::ggsurvplot(KM,pval = TRUE, risk.table = TRUE, ylim = c(0.5,1), main = "kaplan meier estimates of incident main-comorbidities \n adult")
```

## stratified by AED
```{r, fig.height=10, fig.width=10}
u1 <- PATIENTS_AGE_REGION %>% filter(isAED == TRUE) %>% filter(IXDAYS >= -2*365, IXDAYS <= 90) %>% select(ENROLID, SUBSTANCENAME)  %>% distinct()
u2 <- left_join(df.survival, u1)

for (m in MAIN.COMORB) {
KM <- survfit(Surv(time, event) ~ SUBSTANCENAME, type="kaplan-meier", conf.type="log", data=u2 %>% filter(MC == m))
print(
  survminer::ggsurvplot(
  KM,
  ylim = c(0.5, 1),
  risk.table = TRUE, pval = TRUE,
  risk.table.fontsize = 5 , font.legend = 7, risk.table.height = 0.5,
  main = paste("kaplan meier estimates of incident main-comorbidities \n adult",m)
  )
  )
}
```


# treatment difference for incidence patients in medical history


plot only, stratified by MC
```{r, fig.height=10, fig.width=10}
library(data.table)
# sum of patients per main comorbidity
mc.inc180

# number of patients by MC, by AED
GetAEDsPerMC <- function(x) {
  PATIENTS_AGE_REGION %>%
  filter(ENROLID %in% x, IXDAYS >= -2*365, IXDAYS <= 90, isAED == TRUE) %>% # medical history of AED treatment
  dplyr::select(ENROLID, SUBSTANCENAME) %>% distinct() %>% group_by(SUBSTANCENAME) %>% tally()
}

aedsPerMC        <- lapply(MAIN.COMORB, FUN = function(x) GetAEDsPerMC(main.comorb.first.ixday.inc$ENROLID[main.comorb.first.ixday.inc$MAIN.COMORB == x]))
names(aedsPerMC) <- MAIN.COMORB
aedsPerMC        <- do.call(rbind, lapply(aedsPerMC, data.frame, stringsAsFactors=FALSE)) # list to df

# add column with main-comorbidity
MC               <- do.call(rbind, strsplit(row.names(aedsPerMC), split = "\\."))[,1]
MC               <- gsub("Stroke","Stroke.IschemAttack", MC)
row.names(aedsPerMC) <- c()
aedsPerMC$MC         <- MC

# add incident popsize for particular MC
aedsPerMC <- aedsPerMC %>% dplyr::rename(n.treated = n)

# spread and gather again to connect each MC with each AED and set missings to 0 (therefore if NA n.treated will be transformed to 0)
x         <- spread(aedsPerMC, MC, n.treated, fill = 0)
aedsPerMC <- gather(x, key = "MC", value = "n.treated", 2:ncol(x) )
aedsPerMC <- aedsPerMC %>% left_join(mc.inc180)

# get non-treated frequency = n.incident - n.treated
aedsPerMC$n.notTreated <- aedsPerMC$n.incident - aedsPerMC$n.treated
aedsPerMC$n.incident <- c()

# reshape to contingency table
aedsPerMC <- gather(aedsPerMC, key = "cond", "n",c(3,4))
setDT(aedsPerMC)
aedsPerMC.t <- dcast(aedsPerMC, MC + cond ~ SUBSTANCENAME, value.var = "n", fill = 0)


# plot
#######
# reshape aedsPerMC.t for ggplot
x1 <- gather(aedsPerMC.t, "AED","n",3:ncol(aedsPerMC.t))
x2 <- spread(x1, cond, n)
x2$percentTreated <- round( x2$n.treated / x2$n.notTreated, 2) * 100

x2$n.incident <- x2$chisqTest.pval <- c()

ggplot(x2, aes(x = AED, y=percentTreated)) +
  geom_bar(stat="identity") +
  facet_wrap(~ MC,nrow = 3) +
  theme(plot.title = element_text(size=10)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("treatment patterns of incident patients in medical history [-2y,+3months]") +
  xlab("treatment") + ylab("% of incident population") +
  guides(fill=guide_legend(ncol=1)) +
  ylim(c(0,100)) +
  geom_hline(yintercept = 100, linetype = "longdash", colour = "grey") +
  # geom_text(aes(label=paste(percentTreated,"% =",n.treated,"/",n.notTreated)), hjust=-0.5, size = 3) +
  coord_flip()
```


## stratified by AED,  chisq-test and  plot
```{r}
library(data.table)

# sum of patients per main comorbidity
mc.inc180

# number of patients by MC, by AED
GetAEDsPerMC <- function(x) {
  PATIENTS_AGE_REGION %>% 
  filter(ENROLID %in% x, IXDAYS >= -2*365, IXDAYS <= 90, isAED == TRUE) %>% # medical history of AED treatment
  dplyr::select(ENROLID, SUBSTANCENAME) %>% distinct() %>% group_by(SUBSTANCENAME) %>% tally()
}

aedsPerMC        <- lapply(MAIN.COMORB, FUN = function(x) GetAEDsPerMC(main.comorb.first.ixday.inc$ENROLID[main.comorb.first.ixday.inc$MAIN.COMORB == x]))
names(aedsPerMC) <- MAIN.COMORB
aedsPerMC        <- do.call(rbind, lapply(aedsPerMC, data.frame, stringsAsFactors=FALSE)) # list to df

# add column with main-comorbidity
MC               <- do.call(rbind, strsplit(row.names(aedsPerMC), split = "\\."))[,1]
MC               <- gsub("Stroke","Stroke.IschemAttack", MC)
row.names(aedsPerMC) <- c()
aedsPerMC$MC         <- MC

# add incident popsize for particular MC
aedsPerMC <- aedsPerMC %>% dplyr::rename(n.treated = n)

# spread and gather again to connect each MC with each AED and set missings to 0 (therefore if NA n.treated will be transformed to 0)
x         <- spread(aedsPerMC, MC, n.treated, fill = 0)
aedsPerMC <- gather(x, key = "MC", value = "n.treated", 2:ncol(x) )
aedsPerMC <- aedsPerMC %>% left_join(mc.inc180)

# get non-treated frequency = n.incident - n.treated
aedsPerMC$n.notTreated <- aedsPerMC$n.incident - aedsPerMC$n.treated
aedsPerMC$n.incident <- c()

# reshape to contingency table
aedsPerMC <- gather(aedsPerMC, key = "cond", "n",c(3,4))
setDT(aedsPerMC)
aedsPerMC.t <- dcast(aedsPerMC, SUBSTANCENAME + cond ~ MC, value.var = "n", fill = 0)
aedsPerMC.t[is.na(aedsPerMC.t)] <- 0


# chisq test all AEDs
#####################
cat("------------------------------\n")
cat("chisq test all AEDs \n")
s <- unique(aedsPerMC.t$SUBSTANCENAME)
c.test <- c()
for(ss in s){
tmp <- chisq.test(setDF(aedsPerMC.t) %>%  filter(SUBSTANCENAME == ss) %>% dplyr::select(-SUBSTANCENAME,-cond))
c.test <- c(c.test, tmp$p.value)
}

chisq.test.pval <- data.frame(AED=s, chisqTest.pval=c.test)
knitr::kable(chisq.test.pval)

cat("global test \n")
chisq.test(aedsPerMC.t[,3:ncol(aedsPerMC.t)])
```

## plot results
```{r, fig.height=10, fig.width=12}

# reshape aedsPerMC.t for ggplot
x1 <- gather(aedsPerMC.t, "MC","n",3:ncol(aedsPerMC.t))
x2 <- spread(x1, cond, n)
x2$percentTreated <- round( x2$n.treated / x2$n.notTreated, 2) * 100

# bind nPatients to column MC for title
x2 <- x2 %>% left_join(chisq.test.pval, by = c("SUBSTANCENAME" = "AED"))
x2$SUBSTANCENAME <- paste(x2$SUBSTANCENAME, "\n -log10 pval = ", round(-log10(x2$chisqTest.pval),1), sep = "")
x2$n.incident <- x2$chisqTest.pval <- c()

ggplot(x2, aes(x = MC, y=percentTreated)) + 
  geom_bar(stat="identity") + 
  facet_wrap(~ SUBSTANCENAME,nrow = 3) +
  theme(plot.title = element_text(size=10)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("treatment patterns of incident patients in medical history [-2y,+3months]") +
  xlab("treatment") + ylab("% of incident population") +
  guides(fill=guide_legend(ncol=1)) +
  ylim(c(0,100)) + 
  geom_hline(yintercept = 100, linetype = "longdash", colour = "grey") +
  geom_text(aes(label=paste(percentTreated,"% =",n.treated,"/",n.notTreated)), hjust=-0.5, size = 3) +
  coord_flip()
```


# patient frequency by filter criterium
```{r}

ggplot(gather(patObsCount$value()[-2],"age","n",2:3), aes(x=reorder(description,n),y=n, fill=age, label = n)) +
  geom_bar(aes(fill=age), position="dodge",stat="identity") +
  geom_text(aes(label=n), position=position_dodge(0.9), hjust = -0.1) +
  ggtitle("patient frequency by filter criterium and by age") +
  coord_flip() +
  ylim(0,1e6) 

knitr::kable(patObsCount$value())
```


# descriptive stats
* age histogram
* AED histogram
* non-AED histogram, top 20
* diagnosis histogram (by PHEWAS, PHEWAS_highlevel)

*The Truven Health MarketScan® Databases Insurance classes:*

>The Truven Health MarketScan Databases consist of several different datasets. The Commercial Claims and Encounters (CCAE) database is a nationally representative collection of inpatient, outpatient, and pharmaceutical claims from more than 200 insurance carriers and large, self-insuring companies. The Medicare Supplemental and Coordination of Benefits (MDCR) database consists of claims for Medicare-eligible retirees with employer-sponsored supplemental plans.

```{r}
# age 
######

# age histogram
ggplot(PATIENTS_AGE_REGION %>% dplyr::select(ENROLID, AGE) %>% distinct(), aes(AGE)) +
  ggtitle(paste("AGE histogram","\n total pop.-size = ",PATIENTS_AGE_REGION %>% dplyr::select(ENROLID) %>% distinct() %>% tally() )) +
  geom_histogram(aes(y = ..density..))


# THERCLS histogram
#####################

# THERCLS histogram 
ggplot(
    PATIENTS_AGE_REGION %>% 
    filter(SUBSTANCENAME != "", THERCLS != "") %>% 
    dplyr::select(THERCLS, isAED, ENROLID) %>% 
    distinct() %>% 
    group_by(isAED, THERCLS) %>% 
    tally(sort=TRUE) %>% 
    top_n(20,n),
  aes(x = reorder(THERCLS, n), y = n, fill= isAED)
  ) +
  coord_flip() +
  geom_bar(stat="identity") +
  ggtitle(paste("THERCLS frequency, top 2x20","\n total pop.-size = ",PATIENTS_AGE_REGION %>% dplyr::select(ENROLID) %>% distinct() %>% tally() )) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# SUBSTANCE frequencies
ggplot(
  PATIENTS_AGE_REGION %>%
    dplyr::select(ENROLID, isAED, SUBSTANCENAME) %>%
    distinct() %>%
    group_by(isAED, SUBSTANCENAME) %>%
    tally(sort = TRUE) %>%
    top_n(20, n) %>% 
    filter(n >= 100),
  aes(x = reorder(SUBSTANCENAME, n), y = n)) +
  facet_wrap( ~ isAED) +
  coord_flip() +
  geom_bar(stat = "identity") +
  ggtitle(
    paste(
      "SUBSTANCE frequency, top 20",
      "\nnon-AED vs AED",
      "\n total pop.-size = ",
      PATIENTS_AGE_REGION %>% dplyr::select(ENROLID) %>% distinct() %>% tally())
    ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# DIAGNOSIS by PHEWAS_STRING
ggplot(
  PATIENTS_AGE_REGION %>%
    filter(PHEWAS_STRING != "") %>% 
    dplyr::select(MAIN.COMORB, ENROLID, PHEWAS_STRING) %>%
    distinct() %>%
    group_by(MAIN.COMORB, PHEWAS_STRING) %>%
    tally(sort = TRUE) %>%
    top_n(20, n),
  aes(x = reorder(PHEWAS_STRING, n), y = n, fill = MAIN.COMORB)) +
  coord_flip() +
  geom_bar(stat = "identity") +
  ggtitle(
    paste(
      "PHEWAS frequency, top 20 (each category)",
      "\n total pop.-size = ",
      PATIENTS_AGE_REGION %>% dplyr::select(ENROLID) %>% distinct() %>% tally())
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


# status mail
```{r}
source("~/epilepsie_preprocessing_descrStat/functions/StatusEmail.R")
StatusEmail(subject = paste("RStudio status email: ","descr.stat.18to66.Rmd","finished"),body = "x")
```
